---
title: "Group8_Homework2"
format: html
editor: visual
---

## Task 1: Basic Vector Practice

### 1. Creating the Blood Pressure Vectors
Two vectors are created corresponding to the pre-treatment and post-treatment measurements for twenty subjects.

```{r}
# Create the pre-treatment blood pressure vector
pretreatment <- c(130, 128, 116, 124, 133, 134, 118, 126, 114, 127, 141, 138, 128, 140, 137, 131, 120, 128, 139, 135)

# Create the post-treatment blood pressure vector
posttreatment <- c(114, 98, 113, 99, 107, 116, 113, 111, 119, 117, 101, 119, 130, 122, 106, 106, 124, 102, 117, 113)
```

### 2. Assigning Names to the Vector Elements
Subject names are assigned using the paste() function to the pre-treatment and post-treatment vectors.

```{r}
#Assign subject names to pre-treatment vector
names(pretreatment) <- paste("Subject", 1:20, sep = "_")

#Assign subject names to post-treatment vector
names(posttreatment) <- paste("Subject", 1:20, sep = "_")
```

### 3. Calculating Change in Blood Pressure for Each Subject
Calculated the change in blood pressure (pretreatment - posttreatment) and saved the calculation as a new vector.

```{r}
diff_bp <- pretreatment - posttreatment
```

### 4. Calculating Average Decrease For All Subjects
Calculated the average decrease in blood pressure across all subjects.

```{r}
mean(diff_bp)
```
### 5. Determining which Subjects Experienced Decrease in BP with Treatment
Determined a list of subjects that experienced a decrease in blood pressure (diff_bp>0) with treatment.

```{r}
which(diff_bp>0)
```

### 6. Subset Vectors
Subset the vector of differences to only return those that have a positive change.

```{r}
positive_diff <- diff_bp[diff_bp>0]

positive_diff
```
### 7. Calculating Average Decrease for Subset of Subjects
Calculated the average decrease in blood pressure for those subjects that had a positive change 

```{r}
mean(positive_diff)
```

## Task 2: Basic Data Frame Practice

### 1. Creating a Data Frame
Created a data frame object with four columns corresponding to data from Task 1. 

```{r}
# Create a vector containing subject names
patient <- names(pretreatment)

# Combine the blood pressure information into a data frame
bp_dataframe <- data.frame(
    patient = patient,
    pre_bp = pretreatment,
    post_bp = posttreatment,
    diff_bp = diff_bp
)

bp_dataframe
```
### 2. Returning Rows with Select Column Values
Returned only rows where diff_bp column is negative by accessing the column from data frame.

```{r}
bp_dataframe[bp_dataframe$diff_bp<0, ]
```
### 3. Adding a Column to Existing Data Frame
Added a column to the data frame corresponding to "TRUE" if post_bp < 120

```{r}
# Add a TRUE/FALSE column based on post-treatment blood pressure
bp_dataframe$post_lessThan_120 <- bp_dataframe$post_bp < 120

bp_dataframe
```

### 4. Printing Data Frames
Displayed completed blood pressure data frame in a formatted table. 

```{r}
# Display the completed data frame as a formatted table
knitr::kable(bp_dataframe, row.names = FALSE)
```

## Task 3: List Practice

### 1. Creating Placebo Data Frame
Pre-treatment and post-treatment blood pressure measurements from the placebo experiment are used to create a second data frame similar to the treatment data frame.

```{r}
# Create placebo pre-treatment blood pressure vector
placebo_pre <- c(138, 135, 147, 117, 152, 134, 114, 121, 131, 130)

# Create placebo post-treatment blood pressure vector
placebo_post <- c(105, 136, 123, 130, 134, 143, 135, 139, 120, 124)

# Assign subject names to placebo_pre vector
names(placebo_pre) <- paste("PlaceboSubject", 1:10, sep = "_")

# Assign subject names to placebo_post vector
names(placebo_post) <- paste("PlaceboSubject", 1:10, sep = "_")

# Calculate change in blood pressure for placebo
placebo_diff_bp <- placebo_pre - placebo_post

# Create vector containing placebo subject names
placebo_patient <- names(placebo_pre)

#Combine placebo blood pressure information into a data frame
placebo_dataframe <- data.frame(
  patient = placebo_patient,
  pre_bp = placebo_pre,
  post_bp = placebo_post,
  diff_bp = placebo_diff_bp
)

# Add a TRUE/FALSE column based on post-placebo blood pressure
placebo_dataframe$post_lessThan_120 <- placebo_dataframe$post_bp < 120

placebo_dataframe
```

### 2. Create and Store Lists
A list is created to store the treatment and placebo data frames as separate named elements. 

```{r}
# Create a list containing the treatment and placebo data frames
bp_list <- list(
  treatment = bp_dataframe,
  placebo = placebo_dataframe
)

bp_list
```
### 3. Accessing First List Element
The treatment data frame, which is the first element of "bp_list", is accessed using three different list-indexing methods.

```{r}
# Access first element using single brackets
bp_list[1]

# Access first element using double brackets
bp_list[[1]]

# Access first element using its name
bp_list$treatment
```
### 4. Accessing a Specified Column of a Data Frame
The "pre_bp" column is accessed directly from the placebo data frame stored within "bp_list".

```{r}
# Access the placebo pre-treatment blood pressure column
bp_list$placebo$pre_bp
```


## Task 4: Control Flow Practice

