---
title: "Group8_Homework2"
format: html
editor: visual
---

## Task 1: Basic Vector Practice

### 1. Creating the Blood Pressure Vectors
Two vectors are created corresponding to the pre-treatment and post-treatment measurements for twenty subjects.

```{r}
# Create the pre-treatment blood pressure vector
pretreatment <- c(130, 128, 116, 124, 133, 134, 118, 126, 114, 127, 141, 138, 128, 140, 137, 131, 120, 128, 139, 135)

# Create the post-treatment blood pressure vector
posttreatment <- c(114, 98, 113, 99, 107, 116, 113, 111, 119, 117, 101, 119, 130, 122, 106, 106, 124, 102, 117, 113)
```

### 2. Assigning Names to the Vector Elements
Subject numbers are assigned using the paste() function to the pre-treatment and post-treatment vectors.

```{r}
#Assign subject names to pre-treatment vector
names(pretreatment) <- paste("Subject", 1:20, sep = "_")

#Assign subject names to post-treatment vector
names(posttreatment) <- paste("Subject", 1:20, sep = "_")
```

### 3. Calculating Change in Blood Pressure for Each Subject
Calculated the change in blood pressure (pretreatment - posttreatment) and saved the calculation as a new vector.

```{r}
bp_diff <- pretreatment - posttreatment
```

### 4. Calculating Average Decrease For All Subjects
Calculated the average decrease in blood pressure across all subjects.

```{r}
mean(bp_diff)
```
### 5. Determining which Subjects Experienced Decrease in BP with Treatment
Determined a list of subjects that experienced a decrease in blood pressure (bp_diff>0) with treatment.

```{r}
which(bp_diff>0)
```

### 6. Subset Vectors
Subset the vector of differences to only return those that have a positive change.

```{r}
positive_diff <- bp_diff[bp_diff>0]
```
### 7. Calculating Average Decrease for Subset of Subjects
Calculated the average decrease in blood pressure for those subjects that had a positive change 

```{r}
mean(positive_diff)
```


## Task 2: Basic Data Frame Practice

## Task 3: List Practice

## Task 4: Control Flow Practice

