---
title: "Homework2"
author: "Haverhals"
format: html
editor: visual
---

## Task 1

### 1.

First, I am creating vectors with the subject's pre- and post-treatment blood pressure readings.

```{r}
pre<-c(130, 128, 116, 124, 133, 134, 118, 126, 114, 127, 141, 138, 128, 140, 137, 131, 120, 128, 139, 135)
post<-c(114, 98, 113, 99, 107, 116, 113, 111, 119, 117, 101, 119, 130, 122, 106, 106, 124, 102, 117, 113)
```

### 2.

Next, for each set of readings, I'm going to assign names to the values, so we can identify the subjects' measured values moving forward. To do this, I used the paste() function (as recommended) to create a vector of names. Once I had those, I applied these names using the names() function.

```{r}
nms<-paste("subject",1:20,sep="_")
names(pre)<-nms
names(post)<-nms
pre
post
```

### 3.

Next, we were to calculate the differences observed, for each patient. Here, R is subtracting element-wise. It's maybe not worth mentioning, but negative values indicate a lowering of the subjects' blood pressures.

```{r}
delt<-post-pre
delt
```

### 4.

The average change in blood pressure is shown below. As you can see, I used the R function mean() to calculate that number.

```{r}
mean(delt)
```

### 5.

Here, I'm using the which() function to find identify the subject who experienced a drop in blood pressure as a result of the treatment. The R function which() checks the vector an element at a time and finds for us which ones have negative values. Note that only identification is occurring; at this point R is only telling us where the values less than 0 show up (not what the negative vlaues actually are).

```{r}
which(delt<0)
```

### 6.

Focusing on these subjects, here are the individual decreases observed. In other words, R has created a vector out of the subset of the blood pressures (and their associated names) from only those patients who experienced a decrease in blood pressure.

```{r}
drops<-delt[delt<0]
drops
```

### 7.

Finally, the last item in this task asked us to find the average decrease, but only from those subjects who experienced a decrease. Upon getting to this task, I realized I should have naved the vector from #5, so I went back and did that to make the code below a bit cleaner.

```{r}
mean(drops)
```

## Task 2

### 1.

For this problem, I'm going to take the vectors I've defined above and use them to create a data frame.I hope the fact that I have different names here is alright.

```{r}
subject_data<-data.frame(pre,post,delt)
head(subject_data)
```

Note that when I checked this, as you can see, I am getting the line identifiers matching the names. It made things look redundant so I looked around online and found the unname function to clean things up a bit.

```{r}
subject_data<-data.frame(subject=nms,pre_bp=unname(pre),post_bp=unname(post),bp_change=unname(delt))
head(subject_data)
```

Looks better to me.

### 2.

Next, we want to again look at the subset of subjects who experienced a decrease in blood pressure. This time, I used the subset function to get experience with it.

```{r}
drop_frame<-subset(subject_data,delt<0,1:4)
drop_frame
```

### 3.

Here, we're to add a column to the data frame that shows TRUE if the post-treatment blood pressure is less than 120. Using the problem statement as I guide, I'm just going to define that as a vector and hope it gets added to the data frame. Wish me luck!

```{r}
subject_data$post_below_120<-c(subject_data$post_bp<120)
head(subject_data)
```

It worked!

### 4.

Finally, here's a table featuring what we've done so far.

```{r}
knitr::kable(subject_data)
```

## Task 3

### 1.

For this task, we're to create another data frame with patients who recieved a placebo. To do so, I'm going to borrow heavily from the code above but (hopefully) be smarted about how I name things.

```{r}
pl_pre<-c(138, 135, 147, 117, 152, 134, 114, 121, 131, 130)
pl_post<-c(105, 136, 123, 130, 134, 143, 135, 139, 120, 124)
pl_diff<-pl_post-pl_pre
pl_names<-paste("pl subject",1:10,sep="_")
pl_post_below_120<-pl_post<120
pl_data<-data.frame(pl_names,pl_pre,pl_post,pl_diff,pl_post_below_120)
```

### 2.

Next, I'll create a list with two elements which are the data frames from Task 2 and Task 3, respectively.

```{r}
bp_list<-list(treatment=subject_data,placebo=pl_data)
```

### 3.

For this, we need to access the first list element using three different methods. For the first, I'll use the single square bracket option:

```{r}
bp_list[1]
```

Note that this returned the entire data frame, but as a list with a single element.

The next type of syntax uses the double bracket:

```{r}
bp_list[[1]]
```

Here, the element I get back is the data frame itself.

For my ~~final trick~~ type of syntax, I'll refer to the original data frame, but remove the second list element.

```{r}
bp_list[[-2]]
```

### 4.

```{r}
bp_list[[2]]$pl_pre
```

## Task 4

### 1.

Next, I'll add columns to data frames the subjects' post-treatment blood pressures.

```{r}
bp_list[[1]]$status<-character(20)
bp_list[[2]]$status<-character(10)
for (i in 1:20){
  if (bp_list[[1]]$post_bp[i]<=120){
    bp_list[[1]]$status[i]<-"optimal"
  } else if (bp_list[[1]]$post_bp[i]<=130){
    bp_list[[1]]$status[i]<-"borderline"
  } else {
    bp_list[[1]]$status[i]<-"high"
  }
}
for (i in 1:10){
  if (bp_list[[2]]$pl_post[i]<=120){
    bp_list[[2]]$status[i]<-"optimal"
  } else if (bp_list[[2]]$pl_post[i]<=130){
    bp_list[[2]]$status[i]<-"borderline"
  } else {
    bp_list[[2]]$status[i]<-"high"
  }
}
bp_list
```
