---
title: "Homework 3"
author: "Carlos Petzold"
format: html
---

# Task 1: Conceptual Questions

* 1) If your working directory is myfolder/homework/, what relative path would you specify to get the file located at myfolder/MyData.csv?
 
  * If my working directory is myfolder/homework/, the relative path that I specify to get the file located at myfolder/Mydata.csv is ../MyData.csv

* 2) What are the major benefits of using R projects?

  * The major benefits of using R Projects are reproducibility and the ability to easily work with others on the same set of files. Each R project has its own working directory, workspace, history, and folder structure, so by sharing an R project with another person, thet will be able to pick it up and use the paths without any headache. 
  
* 3) What is Git and what is GitHub?

  * Git is a version control software that tracks changes made to files that are being used my multiple people, and GitHub is a web platform where programmers can share and manage their programs/code. 
  
* 4) What are the two main differences between a tibble and a data.frame?

  * A tibble does not print completely in the console when called, but a data frame does. Also, tibbles do not coerce down to a vector when you subset to only 1 column using [].
  
  
# Task 2: Reading Delimited Data

## 1. First, we are reading in the glass.data dataset from the UCI machine learning repository

```{r Task 2, Glass Data}

#We'll use read_csv from the readr package to read in the csv file from the URL:

glass_data <- readr::read_csv("https://www4.stat.ncsu.edu/~online/datasets/glass.data",
                       col_names = c("ID", "refractive_index", "sodium",
                                     "magnesium", "aluminum", "silicon",
                                     "potassium", "calcium", "barium",
                                     "iron", "glass_type"))

glass_data

```


## 2. Next, we are reading in the yeast.data dataset from the UCI machine learning repository

```{r Task 2, Yeast Data}

#We'll use read_tsv from the readr package to read in the tab-seperated file from the URL:

yeast_data <- readr::read_table("https://www4.stat.ncsu.edu/~online/datasets/yeast.data",
                         col_names = c("seq_name", "mcg", "gvh", "alm", "mit", "erl",
                                       "pox", "vac", "nuc", "class"))

yeast_data

```


# Task 3: Combining Excel and Delimited Data

## 3. First we will read in the first sheet of the white wine excel file

```{r Task 3, read Wine data first sheet}

#Reading in first sheet of the file
wine_data <- readxl::read_excel("data/white-wine.xlsx", sheet = 1)

wine_data

```

## 4. Next we will read in the second sheet of the white wine excel file, and assign

```{r Task 3, read Wine data second sheet}

#Reading in second sheet of the file
wine_data_colnames <- readxl::read_excel("data/white-wine.xlsx", sheet = 2)

wine_data_colnames

#Assigning the vector of names from the second sheet to be the column names of the first sheet
colnames(wine_data) <- wine_data_colnames[[1]]

```

## 5. Next we will add a column to indicate if the wines are white (in this case, all are)

```{r Task 3, Wine data white indicator}

#Using dplyr::mutate to add a white wine indicator
wine_data <- wine_data |>
  dplyr::mutate(wine_type = "white")

```

## 6. Next we will read in a red wine dataset (semi-colon delimited)

```{r Task 3, read red wine}

#Using dplyr::mutate to add a white indicator
red_wine_data <- readr::read_delim("https://www4.stat.ncsu.edu/~online/datasets/red-wine.csv",
                                  delim = ";")

#replacing column names
colnames(red_wine_data) <- wine_data_colnames[[1]]

#Using dplyr::mutate to add a red wine indicator
red_wine_data <- red_wine_data |>
  dplyr::mutate(wine_type = "red")

```

## 7. Lastly, we will combine the 2 wine data sets

```{r Task 3, combine wine data}

#combining the 2 datasets using dplyr::bind_rows
wine_data_combined <- dplyr::bind_rows(wine_data, red_wine_data)

```


# Task 4: Writing a Function

## 8. Write a basic function (call it getIQR) that takes in a vector of numbers and returns the interquartile range (75th percentile - 25th percentile). You can use the quantile function. Remove the name on the returned vector (i.e. make sure the returned item doesn’t have a name like 75% or something like that). Allow your function to pass unnamed arguments to the quantile function.

```{r Task 4, IQR function}

#this function takes in a vector and returns the interquartile range of the vector
getIQR <- function(vec = c(1,2,3)) {
  
  if (is.numeric(vec)) {
    
    iqr <- quantile(vec)[[4]] - quantile(vec)[[2]]
    return(unname(iqr))
    
  } else {
    
    return(cat("ERROR: Input vector must be numeric"))
    
  }
  
}

```

## 9. Take the object from question 7 and apply the getIRQ() function to the fixed_acidity and citric_acid columns (two function calls).

```{r Task 4, IQR function calls}

#find IQR of fixed_acidity variable
getIQR(wine_data_combined$fixed_acidity)

#find IQR of citric_acid variable
getIQR(wine_data_combined$citric_acid)

```

## 10. Write another function (call it getCoefVar) that returns the sample coefficient of variation, defined as the sample standard deviation divided by the absolute value of the sample mean. Allow the function to pass unnamed arguments to sd and mean

```{r Task 4, getCoefVar function}

#this function returns the sample coefficient of variation, defined as the sample standard deviation divided by the absolute value of the sample mean
getCoefVar <- function(x, ...) {
  
  samp_coef <- sd(x, ...) / abs(mean(x, ...))
  
  return(samp_coef)
  
}

```

## 11. Repeat question 9 using the getCoefVar() function.

```{r Task 4, getCoefVar function calls}

#find sample coefficient of variation of fixed_acidity variable
getCoefVar(wine_data_combined$fixed_acidity)

#find sample coefficient of variation of citric_acid variable
getCoefVar(wine_data_combined$citric_acid)

```

## 12. Create a function (call it getSummaries) that takes in a vector and returns both the IQR and the coefficient of variation. Do not rewrite your above two functions, call them. When returning your values, give them appropriate names. (This would be called a wrapper function.)

```{r Task 4, getSummaries function}

#this function returns both the IQR and the coefficient of variation
getSummaries <- function(vec) {
  
  output <- c("IQR" = getIQR(vec), "Coef" = getCoefVar(vec))
  
  return(output)
  
}

```

## 13. Repeat question 9 with the getSummaries() function.

```{r Task 4, getSummaries function calls}

#find summary of fixed_acidity variable
getSummaries(wine_data_combined$fixed_acidity)

#find summary of citric_acid variable
getSummaries(wine_data_combined$citric_acid)

```

## 14. Write your own infix function that can be used in replace of the cbind function (feel free to call the cbind function in your function :) Check that the function works on the two data frames below (df1 & df2) and then all three objects (df1, df2, vec1)!

```{r Task 4, infix function and calls}

#this function binds columns
infix <- function(...) {
  cbind(...)
}

#objects for testing
df1 <- data.frame(unif = runif(10), norm = rnorm(10))
df2 <- data.frame(gamma = rgamma(10, shape = 1), beta = rbeta(10, shape1 = 5, shape2 = 1))
vec1 <- rnorm(10, mean = 10)

#function calls
infix(df1,df2)
infix(df1,df2,vec1)
```