---
title: "ST 558 Homework 5"
author: "Group 8: Sheth, Petzold, King, and Haverhals"
format: html
editor: visual
---

## Group Approach

Our individual drafts used several valid approaches to complete the assignment. For the final group submission, we combined those ideas into one consistent workflow. In particular, we used a secure environment variable for the API key, URL encoding so multi-word searches work correctly, explicit parsing of the API response, a compact `across()` approach for factor conversion, and several complementary approaches to categorical exploratory data analysis. The final document uses one consistent set of object names and explanations so that the work reads as a single group submission.

# Task 1: Querying an API and a Tidy-Style Function

## 1. Connecting to the News API

After obtaining a News API key from newsapi.org, the key is stored in an environment variable named `NEWSAPI_KEY` rather than written directly in the homework file. This allows the code to be shared through GitHub without exposing the API key. We use `GET()` from the `httr` package to request recent English-language news articles related to pharmaceuticals and store the response as an R object.

```{r}
#| message: false
#| warning: false

# Load packages used for the API request and parsing
library(httr)
library(jsonlite)
library(tidyverse)

# Retrieve the NewsAPI key from the local environment
news_api_key <- Sys.getenv("NEWSAPI_KEY")

# Query NewsAPI for recent pharmaceutical news
news_response <- GET(
  paste0(
    "https://newsapi.org/v2/everything?q=pharmaceuticals",
    "&from=2026-09-01",
    "&language=en",
    "&apiKey=", news_api_key
  )
)

# Confirm that the request was successful
status_code(news_response)
```

## 2. Parsing the News API Response

Next, we parse the API response and locate the article information. We first convert the returned content to text and parse the JSON. We use `simplifyVector = FALSE` so the nested article information remains in list form. The article records are then combined into a tibble while preserving the nested source information as a list column.

```{r}
# Convert the response content to text and parse the JSON
news_text <- content(news_response, as = "text", encoding = "UTF-8")
news_content <- fromJSON(news_text, simplifyVector = FALSE)

# Extract the article information
article_list <- news_content$articles

# Preserve source as a list column before combining articles
for (i in seq_along(article_list)) {
  article_list[[i]]$source <- list(article_list[[i]]$source)
}

# Combine the article records into a tibble
news_articles <- bind_rows(article_list)

# Inspect the resulting article data
news_articles
```

The resulting object contains the article information returned by the API, and the first column (`source`) is retained as a list column because it contains nested information about each article's source.

## 3. Creating a Function to Query the API

We now create a function that accepts three inputs: a subject to search for, a starting date, and an API key. The search term is URL encoded so that subjects containing spaces can be used. The function sends the API request, parses the returned content, extracts the article information, and returns a tibble.

```{r}
query_news <- function(subject, from_date, api_key) {
  
  # Encode the search term so multi-word subjects can be used
  subject <- URLencode(as.character(subject))
  
  # Build and send the API request
  response <- GET(
    paste0(
      "https://newsapi.org/v2/everything?q=", subject,
      "&from=", from_date,
      "&language=en",
      "&apiKey=", api_key
    )
  )
  
  # Parse the response
  response_text <- content(response, as = "text", encoding = "UTF-8")
  response_content <- fromJSON(response_text, simplifyVector = FALSE)
  article_list <- response_content$articles
  
  # Preserve source as a list column
  for (i in seq_along(article_list)) {
    article_list[[i]]$source <- list(article_list[[i]]$source)
  }
  
  # Return article information as a tibble
  article_data <- bind_rows(article_list)
  return(article_data)
}

# Demonstrate the function with two different searches
pharma_articles <- query_news(
  "pharmaceuticals",
  "2026-09-01",
  news_api_key
)

biotech_articles <- query_news(
  "biotechnology",
  "2026-09-01",
  news_api_key
)

# Display the results
pharma_articles
biotech_articles
```

# Task 2: Summarizing Roller Coaster Data

## Reading in and Preparing the Data

The roller coaster data are read directly from the URL provided in the assignment. Because latitude, longitude, and boundary information are not needed for the EDA in this assignment, those variables are removed. Character variables that describe categorical characteristics are converted to factors using `across()`. `AgeGroup` is converted separately so that its coded values can also be replaced with clearer labels.

```{r}
#| message: false
  
# Read in the roller coaster data
coaster_data <- read_csv(
  "https://www4.stat.ncsu.edu/online/datasets/635USrollercoasters.csv"
)

# Prepare variables for categorical EDA
coaster_data <- coaster_data |>
  select(-Latitude, -Longitude, -Boundary) |>
  mutate(
    across(
      c(Park, City, State, `Census Region`, Scale, Type, Design, `Inversions?`, Make),
      factor
    ),
    AgeGroup = factor(
      AgeGroup,
      levels = c("1:Older", "2:Recent", "3:Newest"),
      labels = c("Older", "Recent", "Newest")
    )
  )
```

# Basic EDA

## 4. Examining How the Data Are Stored

We use `str()` to inspect the structure of the prepared data and verify that the variables are stored appropriately.

```{r}
str(coaster_data)
```

After conversion, the categorical variables are stored as factors, while quantitative measurements such as top speed, maximum height, drop, length, duration, and number of inversions remain numeric. The coaster name remains character data.

## 5. Documenting Missing Values

To document missingness, we use a short function that counts `NA` values within a column and then apply it across all variables. We display only the variables that contain at least one missing value.

```{r}
# Function to count missing values in a column
sum_na <- function(column) {
  sum(is.na(column))
}

# Count missing values for every variable
na_counts <- coaster_data |>
  summarize(across(everything(), sum_na))

# Display only variables containing missing values
na_counts |>
  select(where(~ .x > 0))
```

The data contain missing values in six variables: TopSpeed, MaxHeight, Drop, Length, Duration, and Make. The counts are 74, 8, 128, 33, 66, and 36, respectively, for a total of 345 missing values.

# Categorical Variables

## 6. Creating and Interpreting Contingency Tables

Because exploratory data analysis can examine the same data from several perspectives, we use different categorical combinations for the one-, two-, and three-way tables.

### One-way table

The one-way table summarizes roller coasters by Census Region.

```{r}
table(coaster_data$`Census Region`)
```

For example, the table shows that 134 roller coasters are located in the Midwest region.

### Two-way table

The two-way table compares coaster Scale and Type.

```{r}
table(
  coaster_data$Scale,
  coaster_data$Type
)
```

For example, 216 of the 269 Extreme roller coasters are steel coasters.

### Three-way table

The three-way table compares coaster Type, inversion status, and AgeGroup.

```{r}
table(
  coaster_data$Type,
  coaster_data$`Inversions?`,
  coaster_data$AgeGroup
)
```

For example, 70 roller coasters in the Recent age group are steel coasters that have inversions.

## 7. Creating a Conditional Two-Way Table

Next, we examine the relationship between coaster Scale and Type specifically for coasters in the Older age group. We create the same conditional table in two different ways.

### Method 1: Filter the data first

```{r}
older_coasters <- coaster_data |>
  filter(AgeGroup == "Older")

table(
  older_coasters$Scale,
  older_coasters$Type
)
```

### Method 2: Create a three-way table and subset it

```{r}
scale_type_age <- table(
  coaster_data$Scale,
  coaster_data$Type,
  coaster_data$AgeGroup
)

scale_type_age[, , "Older"]
```

Both methods produce the same conditional two-way table. Using the factor label `"Older"` to subset the three-way table makes the selected category explicit.

## 8. Creating a Two-Way Table with `dplyr`

We next create a two-way contingency table for Scale and Type using `group_by()` and `summarize()`, and then use `pivot_wider()` so the result resembles the output from `table()`.

```{r}
#| message: false

scale_type_summary <- coaster_data |>
  group_by(Scale, Type) |>
  summarize(Count = n())

scale_type_summary |>
  pivot_wider(
    names_from = Type,
    values_from = Count
  )
```

## 9. Creating Stacked and Side-by-Side Bar Graphs

Finally, we visualize the relationship between coaster Scale and Type. The first graph stacks coaster types within each Scale category, while the second places the categories side by side for easier comparison.

### Stacked bar graph

```{r}
ggplot(coaster_data, aes(x = Scale, fill = Type)) +
  geom_bar() +
  labs(
    title = "Roller Coaster Type by Scale: Stacked Bar Plot",
    x = "Roller Coaster Scale",
    y = "Number of Roller Coasters",
    fill = "Coaster Type"
  )
```

### Side-by-side bar graph

```{r}
ggplot(coaster_data, aes(x = Scale, fill = Type)) +
  geom_bar(position = "dodge") +
  labs(
    title = "Roller Coaster Type by Scale: Side-by-Side Bar Plot",
    x = "Roller Coaster Scale",
    y = "Number of Roller Coasters",
    fill = "Coaster Type"
  )
```
