---
title: "Homework_3"
author: "Group 8"
format: html
editor: visual
---

# Task 1: Conceptual Questions

## 1.

1.  **If your working directory is myfolder/homework/, what relative path would you specify to get the file located at myfolder/MyData.csv?** The relative path for this file is "../MyData.csv", where the ".." indicates moving up one directory from the current "homework" folder before accessing "MyData.csv".

2.  **What are the major benefits of using R projects?** R projects help keep analyses organized by giving each project its own working directory, workspace, history, folder structure, and source documents. They also make it easier to use relative file paths, which improves reproducibility and facilitates collaboration by enabling running the same code on different computers/by different users.

3.  **What is Git and what is GitHub?** Git is a version control system that tracks changes made to files over time. It allows users to save versions of their work, review changes, and collaborate without losing previous versions. GitHub is an online platform that hosts Git repositories. It allows users to store their projects remotely, share them with others, collaborate on the same files, and manage changes through tools such as commits, branches, and pull requests.

4.  **What are the two main differences between a tibble and a data.frame?** A tibble is a special type of data frame. It has additional classes (eg, tbl_df and tbl) while still inheriting from data.frame, so it can generally be used anywhere a regular data frame can be used. Additionally, tibble has a different default printing behavior, where when printed it shows the dimensions, variable types, and only rows/columns that fit reasonably on the screen rather than flooding the console with the entire output.

# Task 2: Reading Delimited Data

## 1. Glass Data

First, we'll load in the tidyverse package. Then, we'll reading in the glass data. The raw data is comma delimited, so we use read_csv(), which allows the columns names are set to a vector of the names. Finally, we'll print the tibble.



```{r}
#| warning: false
#| message: false
library(tidyverse)
glassData <- read_csv("https://www4.stat.ncsu.edu/online/datasets/glass.data",col_names = c("ID","RI","NA","MG","AL","SI","K","CA","BA","FE","Type"))

glassData
```

## 2. Yeast Data

Next, we'll read in the yeast data. These raw data is space delimited, so we use read_table and rename the columns, as before, and print the tibble.
```{r}

yeastData <- read_table("https://www4.stat.ncsu.edu/online/datasets/yeast.data",col_names = c("seq_name","mcg","gvh","alm","mit","erl","pox","vac","nuc","class"),show_col_types = FALSE)
yeastData
```

# Task 3: Combining Excel and Delimited Data

## 3. First we will read in the first sheet of the white wine excel file

```{r Task 3, read Wine data first sheet}

#Reading in first sheet of the file
wine_data <- readxl::read_excel("data/white-wine.xlsx", sheet = 1)

wine_data

```

## 4. Next we will read in the second sheet of the white wine excel file, and assign

```{r Task 3, read Wine data second sheet}

#Reading in second sheet of the file
wine_data_colnames <- readxl::read_excel("data/white-wine.xlsx", sheet = 2)

wine_data_colnames

#Assigning the vector of names from the second sheet to be the column names of the first sheet
colnames(wine_data) <- wine_data_colnames[[1]]

```

## 5. Next we will add a column to indicate if the wines are white (in this case, all are)

```{r Task 3, Wine data white indicator}

#Using dplyr::mutate to add a white wine indicator
wine_data <- wine_data |>
  dplyr::mutate(wine_type = "white")

```

## 6. Next we will read in a red wine dataset (semi-colon delimited)

```{r Task 3, read red wine}

#Using dplyr::mutate to add a white indicator
red_wine_data <- readr::read_delim("https://www4.stat.ncsu.edu/~online/datasets/red-wine.csv",
                                  delim = ";")

#replacing column names
colnames(red_wine_data) <- wine_data_colnames[[1]]

#Using dplyr::mutate to add a red wine indicator
red_wine_data <- red_wine_data |>
  dplyr::mutate(wine_type = "red")

```

## 7. Lastly, we will combine the 2 wine data sets and print it out.

```{r Task 3, combine wine data}

#combining the 2 datasets using dplyr::bind_rows
wine_data_combined <- dplyr::bind_rows(wine_data, red_wine_data)
wine_data_combined
```

# Task 4: Writing a <u>Fun</u>ction

## 8.

For this task, we'll start creating functions. To begin, we'll take in a vector and find the IQR.

```{r}
getIQR<-function(x){
  quantile(x,0.75,names=0)-quantile(x,0.25,names=0)
}
```

Note the changing of the names default within the quantiles function so that only the answer is returned. Let's try it!

```{r}
getIQR(c(1:20))
trial_vector<-c(1:100)
getIQR(trial_vector)
```

## 9.

Next, we'll get the interquartile range of the numbers in the fixed_acidity and citric_acid columns of the total_wine tibble above.

```{r}
print("fixed_acidity IQR")
getIQR(wine_data_combined$fixed_acidity)
print("citric_acid IQR")
getIQR(wine_data_combined$citric_acid)
```

## 10.

Next, we'll write a function to get the sample coefficient of variation (defined as the sample standard deviation divided by the absolute value of the sample mean).

```{r}
getCoefVar<-function(x){
  sd(x)/abs(mean(x))
}
```

And now to test:

```{r}
getCoefVar(trial_vector)
getCoefVar(c(1:100))
```

Much success!

## 11.

Now, to apply this function to columns of our tibble.

```{r}
print("fixed_acidity coefficient of variation")
getCoefVar(wine_data_combined$fixed_acidity)
print("citric_acid coefficient of variation")
getCoefVar(wine_data_combined$citric_acid)
```

## 12.

Now, we create a new function to apply the two previous ones.

```{r}
getSummaries<- function(x){
  c("IQR"=getIQR(x),"Coef_of_Var"=getCoefVar(x))
}
```

Giving it a shot:

```{r}
getSummaries(trial_vector)
getSummaries(c(1:100))
```

## 13.

Now, to apply this new function to the columns we've been referring to.

```{r}
print("Summary of fixed_acidity")
getSummaries(wine_data_combined$fixed_acidity)
print("Summary of citric_acid")
getSummaries(wine_data_combined$citric_acid)
```

## 14.

Finally, we'll create a function to replace the cbind function. First, we'll write the function:

```{r}
infix<-function(...){
  cbind(...)
}
```

Now, we'll create the data frames and vector we'll be working on:

```{r}
set.seed(6.28)
df1 <- data.frame(unif = runif(10), norm = rnorm(10))
df2 <- data.frame(gamma = rgamma(10, shape = 1), beta = rbeta(10, shape1 = 5, shape2 = 1))
vec1 <- rnorm(10, mean = 10)
```

Now, to test!

```{r}
infix(df1,df2)
infix(df1,df2,vec1)
```
