---
title: "ST558_Homework5_Sheth"
format: html
editor: visual
---

## Task 1: Querying an API and a Tidy-Style Function

#### 1. Connecting to News API
After obtaining the News API key from newsapi.org, the key was saved to the local console ("NEWSAPI_KEY"). The "httr" package is used to query the News API for recent articles related to pharmaceuticals. The search is limited to English-language articles published since September 1, 2026, and the returned API response is stored as an R object for further processing.

```{r}
# Load package needed to query the API
library(httr)

# Retrieve NewsAPI key from the local RStudio console
news_api_key <- Sys.getenv("NEWSAPI_KEY")

# Query NewsAPI for recent pharmaceutical news
news_response <- GET(
  paste0(
    "https://newsapi.org/v2/everything?q=pharmaceuticals",
    "&from=2026-09-01",
    "&language=en",
    "&apiKey=", news_api_key
  )
)

# Confirm that the API request was successful
status_code(news_response)
```
#### 2. Parsing the News API Response
The response returned by the News API is parsed using "content()". The article information is then extracted from the "articles" element and converted to a tibble for inspection.

```{r}
# Parse the NewsAPI response
news_content <- content(news_response)

# Store the article information as a list
article_list <- news_content$articles
for (i in seq_along(article_list)) {
  article_list[[i]]$source <- list(article_list[[i]]$source)
}

# Combine the individual articles into a tibble
news_article <- dplyr::bind_rows(article_list)

# Display the article data
news_article
```
#### 3. Create a Function to Query API
A function is created to allow a user to query the News API by specifying a search subject, a starting date, and an API key. The function sends the API request, parses the returned content, extracts the article information, and returns the results as a tibble. The function is demonstrated using two different search subjects.

```{r}
# Create a function to query NewsAPI and return article information
query_news <- function(subject, from_date, api_key) {
  response <- GET(
    paste0(
      "https://newsapi.org/v2/everything?q=", subject,
    "&from=", from_date,
    "&language=en",
    "&apiKey=", api_key
    )
  )
response_content <- content(response)
article_list <- response_content$articles
for (i in seq_along(article_list)) {
  article_list[[i]]$source <- list(article_list[[i]]$source)
}
article_data <- dplyr::bind_rows(article_list)
return(article_data)
}

# Using the function to query NewsAPI for pharmaceutical articles
pharma_articles <- query_news("pharmaceuticals", "2026-09-01", news_api_key)

# Display pharmaceutical articles
pharma_articles

# Using the function to query NewsAPI for biotechnology articles
biotech_articles <- query_news("biotechnology", "2026-09-01", news_api_key)

# Display biotech articles
biotech_articles
```

## Task 2: Summarizing Roller Coaster Data

### Reading in Data
The roller coaster data are read from the provided URL. Variables representing categorical characteristics are converted to factors as appropriate. The "AgeGroup" variable is also assigned descriptive factor levels corresponding to the values provided with the data.

```{r}
# Load packages used for data manipulation and visualization
library(tidyverse)

# Read in the roller coaster data
coaster_data <- read_csv("https://www4.stat.ncsu.edu/online/datasets/635USrollercoasters.csv")

# Examine the original structure of the roller coaster data
str(coaster_data)

# Convert categorical variables to factors
coaster_data <- coaster_data |>
  mutate(
    Park = factor(Park),
    City = factor(City),
    State = factor(State),
    `Census Region` = factor(`Census Region`),
    AgeGroup = factor(AgeGroup, 
                      levels = c("1:Older", "2:Recent", "3:Newest"), 
                      labels=c("Older", "Recent", "Newest")
                    ),
   Scale = factor(Scale),
   Type = factor(Type),
   Design = factor(Design),
   `Inversions?` = factor(`Inversions?`),
   Make = factor(Make)
  )  
```

### Basic EDA

#### 4. Examining Data Structure
The structure of the resulting data set from above is then examined to determine whether the variables are stored appropriately. After conversion, the variables appear to be stored appropriately. Categorical variables are stored as factors, while quantitative measurements such as speed, height, drop, length, duration, and number of inversions are stored as numeric variables. The coaster name and boundary information remain character variables, while latitude and longitude remain numeric.

```{r}
# Examine how the roller coaster data are stored
str(coaster_data)
```
#### 5. Documenting Missing Values

### Categorical Variables
