---
title: "ST558_Homework5_King"
format: html
---

# Task 1: Querying an API and a Tidy-Style Function

### 1. Using GET()

```{r}
library(httr)
API_KEY<-"API_KEY"
news_get <- GET(paste0("https://newsapi.org/v2/everything?q=gas&from=2026-09-01&language=en&apiKey=", API_KEY))
```

### 2. Retrieving the Article Info


```{r}
news_content <- content(news_get)
news_articles<- news_content$articles

```

### 3. Function


```{r}
news_grabber <- function(subject,time,api_key){
  GET(paste0("https://newsapi.org/v2/everything?q=",subject,
             "&from=",time,
             "&language=en&apiKey=", api_key))
}

test_grab1<-news_grabber("cats","2026-09-01",API_KEY)
test_grab2<-news_grabber("cats","2026-09-01",API_KEY)

```


# Task 2: Roller Coaster Data

Reading in the coaster data, turning all character variables except for the coaster name and the mapping boundary into a factor.

```{r}
library(tidyverse) 
RC_data <- read.csv("data/635USrollercoasters.csv") |>
  mutate(across(where(is.character) & !Coaster & !Boundary, factor))
```
## Basic EDA

### 1. Data Storage

```{r}
str(RC_data)
```

### Missing Values
```{r}
missing<-is.na(RC_data)
```


### Contingency Tables

One way table:

There are 134 roller coasters in the Midwest region.

```{r}
table(RC_data$Census.Region)
```


Two way table:

There are 4 Thrill-scale roller coasters with a bobsled design.

```{r}
table(RC_data$Design,RC_data$Scale)
```

Three way table:

There are 6 older Extreme-scale roller coasters in the Midwest.

```{r}
table(RC_data$Census.Region,RC_data$Scale,RC_data$AgeGroup)
```







