---
title: "ST558_Homework4_Combined"
author: "Group 8"
format: html
---

# Task 1 : Conceptual Questions

### 1. Rewritten Function Call:

For this question, we're to rewrite the given nested function using R's chaining operator.

The code performs the following: the 'iris' data is first converted to a tibble, variables beginning with "Petal" and the "Species" variable are selected, observations with "Petal.Length \< 1.55" are retained, and the resulting data are sorted by "Species".

```{r}

library(tidyverse)

as_tibble(iris) |>
    select(starts_with("Petal"), Species) |>
    filter(Petal.Length < 1.55) |>
    arrange(Species)

```

### 2. Long versus Wide Data:

Long format data has each row as one observation, and each column as a variable/measurement of that observation. Wide format data may contain more than one observation in a given row, with repeated measurements often spread across multiple columns. For statistical analysis, long format data are generally preferred because most plotting, summarizing, and modeling methods expect data to be in long form.

# Task 2: Reading and Manipulating Data

## Glass Data

### 3. Starting the Chain:

For this problem, we're to replace the integers in the Type_of_glass column of our tibble with the corresponding labels that were given in the problem.

The glass data are read directly from the provided URL and the columns are named.A chain is started on the glass data, and "mutate()" is used to overwrite the numeric "Type_of_glass" variable with a factor containing descriptive labels for each glass type. Code for questions 3 and 4 is presented in the same chunk below.

### 4. Continuing the Chain:

The chain is continued to retain only observations with "Fe" values less than 0.2 and glass types of either "tableware" or "headlamps". Code for questions 3 and 4 is presented in the same chunk below.

```{r}
# Reading in the data from the previous assignment:

glass_names <- c("Id", "RI", "Na", "Mg", "Al", "Si", "K", "Ca", "Ba", "Fe", "Type_of_glass")
glass_data <- read_csv("https://www4.stat.ncsu.edu/~online/datasets/glass.data", 
                       col_names = glass_names)

# Creating the factor variable: 

glass_data_filtered <- glass_data |>
  mutate(
    Type_of_glass = factor(
      Type_of_glass,
      levels = c (1, 2, 3, 4, 5, 6, 7),
      labels = c (
        "building_windows_float_processed", 
        "building_windows_non_float_processed",
        "vehicle_windows_float_processed",
        "vehicle_windows_non_float_processed",
        "containers",
        "tableware",
        "headlamps"
      )
    )
  ) |>
  filter(
    Fe < 0.2,
    Type_of_glass %in% c("tableware", "headlamps")
  )

```

## Yeast Data

### 5. Starting the Chain

First, we'll read in the yeast data from before. Then, a chain is started on the yeast data, and "select()" is used to remove the "seq_name" and "nuc" variables.

### 6. Continuing the Chain

The chain is continued by grouping the observations by "class". The "across()" function within "mutate()" is then used to add the mean and median of each numeric variable ("mcg", "gvh", "alm", "mit", "erl", "pox", and "vac") for each class group. The code for questions 5 and 6 is presented in the same chunk below.

```{r}
# Reading in the data from the previous assignment:

yeast_names <- c("seq_name","mcg","gvh","alm","mit","erl","pox","vac","nuc","class")
yeast_data <- read_table("https://www4.stat.ncsu.edu/online/datasets/yeast.data",col_names = yeast_names)

# Creating the chain:

yeast_data_summary <- yeast_data |>
  select(-"seq_name", -"nuc") |>
  group_by(class)|>
  mutate(across(where(is.numeric),list(mean=mean,median=median),.names="{.col}_{.fn}"))

```

## Wine Data

First, we'll read in the dataset using the code from the previous assignment:

```{r}
# Loading necessary package to read excel data:

library(readxl)

# Reading in the data:

white_wine <- read_excel("data/white-wine.xlsx",col_types = "numeric")
wine_names <-read_excel("data/white-wine.xlsx", sheet = excel_sheets("data/white-wine.xlsx")[2])
colnames(white_wine) <- wine_names$Variables
white_wine$type <- "white"
red_wine <- read_delim("https://www4.stat.ncsu.edu/~online/datasets/red-wine.csv",delim = ";")
colnames(red_wine) <- wine_names$Variables
red_wine$type <- "red"
wine_data <- bind_rows(white_wine,red_wine)
```

### 7. Filtering for High-Quality

A chain is started on the combined wine data, and "filter()" is used to retain only high-quality wines with "quality \> 6.5" and wines with "alcohol \< 132". Code for questions 7 through 10 is presented together in the chunk below.

### 8. Sorting by Quality

The chain is continued using "arrange()" with "desc(quality)" to sort the filtered wine data from highest quality to lowest quality.

### 9. Selecting Variables

The chain is continued using "select()" to retain variables containing "acid", along with the "alcohol", "type", and "quality" variables. Code for questions 7 through 10 is presented together in the chunk below.

### 10. Adding Mean and SD

The chain is continued by grouping the wine data by "quality". The mean and standard deviation of "alcohol" are then calculated for each quality group and added as new variables using "mutate()".

```{r}

# Filtering the data:

wine_data_filtered <- wine_data |>
  filter (
    quality > 6.5,
    alcohol < 13.2 ) |>
  arrange (desc(quality)) |>
  select (contains ("acid"), alcohol, type, quality) |>
  group_by (quality) |>
  mutate (
    alcohol_mean = mean(alcohol),
    alcohol_sd = sd(alcohol))

# Printing the table:

wine_data_filtered

```

# Task 3: Database Practice

## Questions 11-13

### 11. Connect to the Database

A connection is established to the "Lahman.db" database. The available tables in the database are then displayed to identify the data available for the remaining questions.

```{r}
# Load packages needed to connect to SQLite database
library (DBI)
library (RSQLite)

# Connect to the Lahman SQLite database
con <- dbConnect (
  SQLite(),
  "data/Lahman.db")

# Display all tables in the database
dbListTables (con)
```

### 12. Using dplyr for Teams table

The "Teams" table is accessed using "tbl()", and "dplyr" functions are used to return the team data for the year 2015.

```{r}
# Accessing "Teams" table and return data for specific year using dplyr
teams_2015_dplyr <- con |>
  tbl("Teams") |>
  filter(yearID == 2015)

# Display the table
teams_2015_dplyr
```

### 13. Using SQL for Teams table

The 2015 "Teams" data are retrieved again, this time by using an SQL statement within "sql()" and "tbl()". This provides the same requested data using SQL rather than the "dplyr" filtering approach.

```{r}

# Return teams data for the year 2015 using SQL
teams_2015_sql <- tbl(con, sql(
"SELECT *
FROM `Teams`
WHERE (`yearID` = 2015.0)")
)

# Display the table
teams_2015_sql

```

## Questions 14-17

### 14. Hall of Fame Players

The "HallOfFame" table is used to identify players who were inducted into the Hall of Fame. The resulting data retain the player identifier, the year the player was voted in, and the category.

```{r}

# Return Hall of Fame inductees, induction year, and category

hall_of_fame <- tbl (con, "HallOfFame") |>
  filter(inducted == "Y") |>
  select(playerID, yearID, category)

# Display the table
hall_of_fame

```

### 15. Hall of Fame Names

The Hall of Fame results from Question 14 are combined with the "People" table to add each inducted player's first and last name.

```{r}

# Select names from the People table

people_names <- tbl(con, "People") |>
  select(playerID, nameFirst, nameLast)

# Joining the hall_of_fame table with the names

hall_of_fame_names <- hall_of_fame |>
  left_join(
    people_names,
    by = "playerID"
  )

# Display the new table

hall_of_fame_names

```

### 16. Managers Career Summary

The "Managers" table is used to obtain each manager's player identifier, games managed, wins, and losses. The data are grouped by manager and summarized to calculate career totals for games managed, wins, and losses. After collecting the results into R, a career win/loss percentage is calculated as "Total_W/G_managed", and the managers are sorted from highest to lowest percentage.

```{r}

# Creating a chain to manipulate the manager data 

manager_summary <- tbl(con, "Managers") |>
  select(playerID, G, W, L) |>
  group_by(playerID) |>
  summarize(
    G_managed = sum(G, na.rm = TRUE),
    Total_W = sum(W, na.rm = TRUE),
    Total_L = sum(L, na.rm = TRUE)) |>
  collect() |>
  mutate(Win_Loss_Percentage = Total_W / G_managed) |>
  arrange(desc(Win_Loss_Percentage))

# Display the table

manager_summary

```

### 17. Hall of Fame Managers

The Hall of Fame results from Question 15 are combined with the manager career results from Question 16. Only individuals who both managed and were inducted into the Hall of Fame are retained, regardless of Hall of Fame category. All variables from the two results are retained, producing a final data set with nine columns.

```{r}

# Combine Hall of Fame data with Manager Summary data

hall_of_fame_managers <- hall_of_fame_names |>
  collect() |>
  inner_join(
    manager_summary,
    by = "playerID")

# Display Table

```
