---
title: "ST558_Homework4_Sheth"
format: html
editor: visual
---

## Task 1: Conceptual Questions 

#### 1. BaseR's Chaining Operator

The following nested function call is rewritten using baseR's chaining operator ("|>"). The 'iris' data are first converted to a tibble, variables beginning with "Petal" and the "Species" variable are selected, observations with "Petal.Length < 1.55" are retained, and the resulting data are sorted by "Species".
  library(tidyverse)
  arrange(filter(select(as_tibble(iris), starts_with("Petal"), Species), Petal.Length < 1.55), Species)
  
```{r}
# Load tidyverse library
library(tidyverse)

# Convert iris to tibble, select variables, filter observations, and sort by "Species"
as_tibble(iris) |>
    select(starts_with("Petal"), Species) |>
    filter(Petal.Length < 1.55) |>
    arrange(Species)
```

#### 2. Long versus Wide Format Data

Long format data have one observation per row and one variable per column. Wide format data may contain more than one observation in a given row, with repeated measurements often spread across multiple columns.

For statistical analysis, long format data are generally preferred because most plotting, summarizing, and modeling methods expect data to be in long form.

## Task 2: Data Manipulation

### Glass Data

#### 3. Creating a Factor for Type of Glass

The glass data are read directly from the provided URL and the columns are named. A chain is started on the glass data, and "mutate()" is used to overwrite the numeric "Type_of_glass" variable with a factor containing descriptive labels for each glass type. Code for questions 3 and 4 is presented in the same chunk below. 

#### 4. Filtering the Glass Data

The chain is continued to retain only observations with "Fe" values less than 0.2 and glass types of either "tableware" or "headlamps". Code for questions 3 and 4 is presented in the same chunk below. 

```{r}
# Assign column names for data set
glass_names <- c("Id", "RI", "Na", "Mg", "Al", "Si", "K", "Ca", "Ba", "Fe", "Type_of_glass")

# Read in comma delimited data from URL
glass_data <- read_csv("https://www4.stat.ncsu.edu/~online/datasets/glass.data", 
                       col_names = glass_names
                      )

# Replace numeric glass type values with descriptive factor labels
glass_data_filtered <- glass_data |>
  mutate(
    Type_of_glass = factor(
      Type_of_glass,
      levels = c (1, 2, 3, 4, 5, 6, 7),
      labels = c (
        "building_windows_float_processed", 
        "building_windows_non_float_processed",
        "vehicle_windows_float_processed",
        "vehicle_windows_non_float_processed",
        "containers",
        "tableware",
        "headlamps"
      )
    )
  ) |>
  filter(
    Fe < 0.2,
    Type_of_glass %in% c("tableware", "headlamps")
  )

# Print the resulting tibble
glass_data_filtered
```
### Yeast Data

#### 5. Removing Variables from Data Set

The yeast data are read directly from the provided URL and column names are assigned. A chain is started on the yeast data, and "select()" is used to remove the "seq_name" and "nuc" variables. The code for questions 5 and 6 is presented in the same chunk below.

#### 6. Adding Class-Specific Means and Medians

The chain is continued by grouping the observations by "class". The "across()" function within "mutate()" is then used to add the mean and median of each numeric variable ("mcg", "gvh", "alm", "mit", "erl", "pox", and "vac") for each class group. The code for questions 5 and 6 is presented in the same chunk below.

```{r}
# Assign column names for the data set
yeast_names <- c("seq_name", "mcg", "gvh", "alm", "mit", "erl", "pox", "vac", "nuc", "class")

# Read in space (whitespace) delimited data from URL
yeast_data <- read_table("https://www4.stat.ncsu.edu/~online/datasets/yeast.data",
                         col_names = yeast_names
                         )

# Remove variables("seq_name" and "nuc") and calculate class-specific means and medians
yeast_data_summary <- yeast_data |>
  select(-seq_name, -nuc) |>
  group_by(class) |>
  mutate (
    across (
      c (mcg, gvh, alm, mit, erl, pox, vac),
      list (mean = mean, median = median)
    )
  )

# Print the resulting tibble
yeast_data_summary
```

### Wine Data

#### 7. Filtering the Wine Data

The white and red wine data are read and combined as in Homework 3. A chain is started on the combined wine data, and "filter()" is used to retain only high-quality wines with "quality > 6.5" and wines with "alcohol < 132". Code for questions 7 through 10 is presented together in the chunk below. 

#### 8. Sorting the Wine Data

The chain is continued using "arrange()" with "desc(quality)" to sort the filtered wine data from highest quality to lowest quality. Code for questions 7 through 10 is presented together in the chunk below. 

#### 9. Selecting Wine Variables

The chain is continued using "select()" to retain variables containing "acid", along with the "alcohol", "type", and "quality" variables. Code for questions 7 through 10 is presented together in the chunk below. 

#### 10. Adding Alcohol Summary Statistics by Quality

The chain is continued by grouping the wine data by "quality". The mean and standard deviation of "alcohol" are then calculated for each quality group and added as new variables using "mutate()". Code for questions 7 through 10 is presented together in the chunk below. 

```{r}
# Load readxl
library(readxl)

# Read the first sheet of the white wine Excel file
white_wine <- read_excel("data/white-wine.xlsx",
                         sheet = 1
                         )

# Read in the second sheet containing alternative variable names
wine_names <- read_excel("data/white-wine.xlsx",
                         sheet = 2
                         )

# Replace the current white wine column names
colnames(white_wine) <- wine_names[[1]]

# Add a column identifying the wine type
white_wine$type <- "white"

# Read in semicolon delimited data
red_wine <- read_delim("https://www4.stat.ncsu.edu/~online/datasets/red-wine.csv",
                       delim = ";"
                       )

# Replace the red wine column names
colnames(red_wine) <- wine_names[[1]]

# Add a column identifying the wine type
red_wine$type <- "red"

# Combine the white and red wine data sets
wine_data <- bind_rows(white_wine, red_wine)

# Filter high-quality wines, sort from highest to lowest quality, select variables, and calculate alcohol summary statistics by quality
wine_data_filtered <- wine_data |>
  filter (
    quality > 6.5,
    alcohol < 132
  ) |>
  arrange (desc(quality)) |>
  select (contains ("acid"), alcohol, type, quality) |>
  group_by (quality) |>
  mutate (
    alcohol_mean = mean(alcohol),
    alcohol_sd = sd(alcohol)
  )

# Print the resulting tibble
wine_data_filtered
```

## Task 3: Database Practice

#### 11. Connecting to the Lahman Database

A connection is established to the "Lahman.db" database. The available tables in the database are then displayed to identify the data available for the remaining questions. (Note, "RSQLite" package installed via Console using "install.packages("RSQLite").)

```{r}
# Load packages needed to connect to SQLite database
library (DBI)
library (RSQLite)

# Connect to the Lahman SQLite database
con <- dbConnect (
  SQLite(),
  "data/Lahman.db"
)

# Display all tables in the database
dbListTables (con)
```
#### 12. Retrieving the 2015 Teams Data Using dplyr

The "Teams" table is accessed using "tbl()", and "dplyr" functions are used to return the team data for the year 2015. 

```{r}
# Accessing "Teams" table and return data for specific year using dplyr
teams_2015_dplyr <- tbl(con, "Teams") |>
  filter(yearID == 2015)

# Display resulting table
teams_2015_dplyr
```

#### 13. Retrieving the 2015 Teams Data Using SQL

The 2015 "Teams" data are retrieved again, this time by using an SQL statement within "sql()" and "tbl()". This provides the same requested data using SQL rather than the "dplyr" filtering approach. 

```{r}
# Return teams data for the year 2015 using SQL
teams_2015_sql <- tbl(
  con,
  sql("select * from teams where yearID = 2015")
)

# Display the resulting table
teams_2015_sql
```

#### 14. Identifying Hall of Fame Players

The "HallOfFame" table is used to identify players who were inducted into the Hall of Fame. The resulting data retain the player identifier, the year the player was voted in, and the category. 

```{r}
# Return Hall of Fame inductees, induction year, and category
hall_of_fame <- tbl (con, "HallOfFame") |>
  filter(inducted == "Y") |>
  select(playerID, yearID, category)

# Display the resulting table
hall_of_fame
```

#### 15. Adding Player Names to the Hall of Fame Data

The Hall of Fame results from Question 14 are combined with the "People" table to add each inducted player's first and last name.

```{r}
# Select player names from the People table
people_names <- tbl(con, "People") |>
  select(playerID, nameFirst, nameLast)

# Add first and last names to the Hall of Fame data
hall_of_fame_names <- hall_of_fame |>
  left_join(
    people_names,
    by = "playerID"
  )

# Display the resulting table
hall_of_fame_names
```

#### 16. Summarizing Manager Career Records

The "Managers" table is used to obtain each manager's player identifier, games managed, wins, and losses. The data are grouped by manager and summarized to calculate career totals for games managed, wins, and losses. After collecting the results into R, a career win/loss percentage is calculated as "Total_W/G_managed", and the managers are sorted from highest to lowest percentage.

```{r}
# Summarize career games, wins, and losses for each manager
manager_summary <- tbl(con, "Managers") |>
  select(playerID, G, W, L) |>
  group_by(playerID) |>
  summarize(
    G_managed = sum(G, na.rm = TRUE),
    Total_W = sum(W, na.rm = TRUE),
    Total_L = sum(L, na.rm = TRUE)
  ) |>
  collect() |>
  mutate(Win_Loss_Percentage = Total_W / G_managed) |>
  arrange(desc(Win_Loss_Percentage))

# Display the resulting table
manager_summary
```

#### 17. Combining Hall of Fame and Manager Information

The Hall of Fame results from Question 15 are combined with the manager career results from Question 16. Only individuals who both managed and were inducted into the Hall of Fame are retained, regardless of Hall of Fame category. All variables from the two results are retained, producing a final data set with nine columns.

```{r}
# Combine Hall of Fame inductees with manager career records
hall_of_fame_managers <- hall_of_fame_names |>
  collect() |>
  inner_join(
    manager_summary,
    by = "playerID"
  )

# Display the resulting table
hall_of_fame_managers
```
