---
title: "Homework 4"
author: "Carlos Petzold"
format: html
---

## Task 1: Conceptual Questions

### 1) Rewrite the following nested function call using baseR’s chaining operator:

library(tidyverse)
arrange(filter(select(as_tibble(iris), starts_with("Petal"), Species), Petal.Length <
1.55), Species)


```{r Task 1 Question 1}

library(tidyverse)

iris <- as_tibble(iris) |> 
  select(starts_with("Petal"), Species) |>
  filter(Petal.Length < 1.55) |>
  arrange(Species)

```


### 2) What is meant by long format data and wide format data? Which do we generally prefer for statistical analysis?

Long data format refers to data that is formatted with 1 observation per row and each column corresponds to one variable. Wide data format refers to data that may have more than one observation for a given row. When conductin statistical analysis, we typically prefer data in the long format because it makes plotting, summarizing, modeling, and analyzing data much simpler when we have one observation per row.


## Task 2: Read in Datasets

### 3) Glass Data: Start a chain on your data object that overwrites the Type_of_glass variable using mutate().

#### Create a factor version of the variable that replaces the numbers with string labels (that is, replace 1 with “building_windows_float_processed”, 2 with “building_win. . . ”, etc. - see the variable descriptions above to give meaningful values).

### 4) Glass Data: Continue your chain and keep only observations where the Fe variable is less than 0.2 and the Type of Glass is either “tableware” or “headlamp”.

```{r Task 2 Question 3 and 4}

#reading in glass data as a csv
glass_data <- readr::read_csv("https://www4.stat.ncsu.edu/~online/datasets/glass.data",
                       col_names = c("ID", "refractive_index", "sodium",
                                     "magnesium", "aluminum", "silicon",
                                     "potassium", "calcium", "barium",
                                     "Fe", "type_of_glass")) |>
  mutate(type_of_glass = factor(
    glass_type,
    levels = c(1, 2, 3, 4, 5, 6, 7),
    labels = c(
      "building_windows_float_processed",
      "building_windows_non_float_processed",
      "vehicle_windows_float_processed",
      "vehicle_windows_non_float_processed",
      "containers",
      "tableware",
      "headlamps"
    )
  )) |>
  filter(Fe < 0.2, type_of_glass %in% c("tableware", "headlamp"))
  
glass_data

```


### 5) Yeast Data: Start a chain on your data object that overwrites the Type_of_glass variable using mutate().

### 6) Yeast Data: Continue your chain to add columns corresponding to the mean and median of each numeric variable (mcg, gvh, alm, mit, erl, pox, and vac) at each class grouping (see the across() function as we did in the dplyr video!).

```{r Task 2 Question 5 and 6}

#reading in yeast data using read_table
yeast_data <- readr::read_table("https://www4.stat.ncsu.edu/~online/datasets/yeast.data",
                         col_names = c("seq_name", "mcg", "gvh", "alm", "mit", "erl",
                                       "pox", "vac", "nuc", "class")) |>
  select(-seq_name, -nuc) |>
  group_by(class) |>
  mutate(across(where(is.numeric), \(x) mean(x), .names = "mean_{.col}"),
         across(where(is.numeric), \(x) median(x), .names = "median_{.col}"))

yeast_data

```


### 7) Wine Data: Start a chain on your new combined data object to filter the data to only look at high-quality wines (quality > 6.5) and wines that have a reasonable alcohol value (alcohol < 13.2).

### 8) Wine Data: Continue your chain to now sort the data from highest quality to lowest.

### 9) Wine Data: Continue your chain to select only the variables that contain acid, the alcohol variable, the type variable, and the quality variable

### 10) Wine Data: Continue your chain to add the mean and standard deviation of the alcohol variable to the data set for each setting of the quality variable.

```{r Task 2 Question 7, 8, 9, and 10}

#Reading in first and second sheet of the file
white_wine_data <- readxl::read_excel("data/white-wine.xlsx", sheet = 1)
wine_data_colnames <- readxl::read_excel("data/white-wine.xlsx", sheet = 2)

#Assigning the vector of names from the second sheet to be the column names of the first sheet
colnames(white_wine_data) <- wine_data_colnames[[1]]

#Using dplyr::mutate to add a white wine indicator
white_wine_data <- white_wine_data |>
  dplyr::mutate(type = "white")

#Reading in red wine data
red_wine_data <- readr::read_delim("https://www4.stat.ncsu.edu/~online/datasets/red-wine.csv",
                                  delim = ";")

#replacing column names
colnames(red_wine_data) <- wine_data_colnames[[1]]

#Using dplyr::mutate to add a red wine indicator
red_wine_data <- red_wine_data |>
  dplyr::mutate(type = "red")

#combining the 2 datasets using dplyr::bind_rows
wine_data_combined <- dplyr::bind_rows(white_wine_data, red_wine_data)

wine_data <- wine_data_combined |>
  filter(quality > 6.5, alcohol < 13.2) |>
  arrange(desc(quality)) |>
  select(fixed_acidity, volatile_acidity, citric_acid, alcohol, type, quality) |>
  group_by(quality) |>
  mutate(mean_alcohol = mean(alcohol),
         sd_alcohol = sd(alcohol))
  

```






