---
title: "ST558_Homework5_Sheth"
format: html
editor: visual
---

## Task 1: Querying an API and a Tidy-Style Function

#### 1. Connecting to News API
After obtaining the News API key from newsapi.org, the key was stored in an environment variable ("NEWSAPI_KEY") through the local RStudio Console. The "httr" package is used to query the News API for recent articles related to pharmaceuticals. The search is limited to English-language articles published since September 1, 2026, and the returned API response is stored as an R object for further processing.

```{r}
# Load package needed to query the API
library(httr)

# Retrieve NewsAPI key from the local RStudio console
news_api_key <- Sys.getenv("NEWSAPI_KEY")

# Query NewsAPI for recent pharmaceutical news
news_response <- GET(
  paste0(
    "https://newsapi.org/v2/everything?q=pharmaceuticals",
    "&from=2026-09-01",
    "&language=en",
    "&apiKey=", news_api_key
  )
)

# Confirm that the API request was successful
status_code(news_response)
```
#### 2. Parsing the News API Response
The response returned by the News API is parsed using "content()". The article information is then extracted from the "articles" element and converted to a tibble for inspection.

```{r}
# Parse the NewsAPI response
news_content <- content(news_response)

# Store the article information as a list
article_list <- news_content$articles
for (i in seq_along(article_list)) {
  article_list[[i]]$source <- list(article_list[[i]]$source)
}

# Combine the individual articles into a tibble
news_article <- dplyr::bind_rows(article_list)

# Display the article data
news_article
```
#### 3. Create a Function to Query API
A function is created to allow a user to query the News API by specifying a search subject, a starting date, and an API key. The function sends the API request, parses the returned content, extracts the article information, and returns the results as a tibble. The function is demonstrated using two different search subjects.

```{r}
# Create a function to query NewsAPI and return article information
query_news <- function(subject, from_date, api_key) {
  response <- GET(
    paste0(
      "https://newsapi.org/v2/everything?q=", subject,
    "&from=", from_date,
    "&language=en",
    "&apiKey=", api_key
    )
  )
response_content <- content(response)
article_list <- response_content$articles
for (i in seq_along(article_list)) {
  article_list[[i]]$source <- list(article_list[[i]]$source)
}
article_data <- dplyr::bind_rows(article_list)
return(article_data)
}

# Using the function to query NewsAPI for pharmaceutical articles
pharma_articles <- query_news("pharmaceuticals", "2026-09-01", news_api_key)

# Display pharmaceutical articles
pharma_articles

# Using the function to query NewsAPI for biotechnology articles
biotech_articles <- query_news("biotechnology", "2026-09-01", news_api_key)

# Display biotech articles
biotech_articles
```

## Task 2: Summarizing Roller Coaster Data

### Reading in Data
The roller coaster data are read from the provided URL. Variables representing categorical characteristics are converted to factors as appropriate. The "AgeGroup" variable is also assigned descriptive factor levels corresponding to the values provided with the data.

```{r}
# Load packages used for data manipulation and visualization
library(tidyverse)

# Read in the roller coaster data
coaster_data <- read_csv("https://www4.stat.ncsu.edu/online/datasets/635USrollercoasters.csv")

# Examine the original structure of the roller coaster data
str(coaster_data)

# Convert categorical variables to factors
coaster_data <- coaster_data |>
  mutate(
    Park = factor(Park),
    City = factor(City),
    State = factor(State),
    `Census Region` = factor(`Census Region`),
    AgeGroup = factor(AgeGroup, 
                      levels = c("1:Older", "2:Recent", "3:Newest"), 
                      labels=c("Older", "Recent", "Newest")
                    ),
   Scale = factor(Scale),
   Type = factor(Type),
   Design = factor(Design),
   `Inversions?` = factor(`Inversions?`),
   Make = factor(Make)
  )  
```

### Basic EDA

#### 4. Examining Data Structure
The structure of the resulting data set from above is then examined to determine whether the variables are stored appropriately. 

After conversion, the variables appear to be stored appropriately. Categorical variables are stored as factors, while quantitative measurements such as speed, height, drop, length, duration, and number of inversions are stored as numeric variables. The coaster name and boundary information remain character variables, while latitude and longitude remain numeric.

```{r}
# Examine how the roller coaster data are stored
str(coaster_data)
```
#### 5. Documenting Missing Values
Missing values are examined for each variable in the roller coaster data. The number of missing observations is calculated for each variable, and only variables containing one or more missing values are displayed. 

Missing values were identified in six variables. There are 74 missing values for TopSpeed, 8 for MaxHeight, 128 for Drop, 33 for Length, 66 for Duration, and 36 for Make. No missing values were identified in the remaining variables.

```{r}
# Count missing values for each variable
missing_values <- colSums(is.na(coaster_data))

# Display only variables that contain missing values
missing_values[missing_values > 0]
```

### Categorical Variables

#### 6. Creating and Interpreting Contingency Tables
Contingency tables are created to summarize the categorical variables in the roller coaster data. A one-way table is first used to examine coaster type. Then a two-way table is used to examine coaster type and inversions. Finally a three-way table is used to examine the previously mentioned variables and age group.

For the one-way contingency table, 530 of the roller coasters are steel coasters. For the two-way contingency table, 166 steel roller coasters have inversions. For the three-way contingency table, 70 roller coasters in the Recent age group are steel coasters that have inversions.

```{r}
# Create a one-way contingency table for coaster type
table(coaster_data$Type)

# Create a two-way contingency table for coaster type and inversions
table(
  coaster_data$Type,
  coaster_data$`Inversions?`
)

# Create a three-way contingency table for coaster type, inversions, and age group
table(
  coaster_data$Type,
  coaster_data$`Inversions?`,
  coaster_data$AgeGroup
)
```
#### 7. Creating Conditional Two-Way Tables
A conditional two-way contingency table is created for coaster type and inversion status, conditioned on the Recent age group. The table is produced first by filtering the original data and then by subsetting a three-way contingency table. Both methods produce the same results.

```{r}
# Method 1: Filter to Recent coasters, then create a two-way table
recent_coasters <- coaster_data |>
  filter(AgeGroup == "Recent")

table(
  recent_coasters$Type,
  recent_coasters$`Inversions?`
)

# Method 2: Create a three-way table
coaster_table <- table(
  coaster_data$Type,
  coaster_data$`Inversions?`,
  coaster_data$AgeGroup
)

coaster_table[, , "Recent"]
```
#### 8. Creating a Two-Way Contingency Table
A two-way contingency table for coaster type and inversion status is created using "group_by()" and "summarize()". The summarized data are then reshaped using "pivot_wider()" so that the output resembles the two-way contingency table created previously with "table()".

```{r}
# Create a two-way contingency table using group_by() and summarize()
coaster_summary <- coaster_data |>
  group_by(Type, `Inversions?`) |>
  summarize(Count=n())

coaster_summary

# Reshape the summarized data to resemble a contingency table
coaster_summary |>
  pivot_wider(
    names_from = `Inversions?`,
    values_from = Count
  )
```
#### 9. 
A stacked bar graph and a side-by-side bar graph are created to visually compare coaster type and inversion status. The stacked bar graph shows the inversion categories within each coaster type, while the side-by-side graph places the categories next to one another for easier comparison.

```{r}
# Create a stacked bar graph of coaster type by inversion status
ggplot(coaster_data, aes(x = Type, fill = `Inversions?`)) +
  geom_bar() +
  labs(
    title = "Stacked Bar Graph of Roller Coaster Type by Inversion Status",
    x = "Coaster Type",
    y = "Number of Roller Coasters",
    fill = "Inversions?"
  )

# Create a side-by-side bar graph of coaster type by inversion status
ggplot(coaster_data, aes(x = Type, fill = `Inversions?`)) +
  geom_bar(position = "dodge") +
  labs(
    title = "Side-by-Side Bar Graph of Roller Coaster Type by Inversion Status",
    x = "Coaster Type",
    y = "Number of Roller Coasters",
    fill = "Inversions?"
  )
```

