---
title: "ST558_Homework4_King"
format: html
---

# Task 1: Conceptual Questions

1. Rewritten nesting call:

```{r}
library(tidyverse)
iris |>
  as_tibble() |>
  select(starts_with("Petal"), Species) |>
  filter(Petal.Length < 1.55) |>
  arrange(Species)
```

2. Long format data has each row as one observation, and each column as a variable/measurement of that observation. Wide format data has more than one observation per row. Long format is preferred.

# Task 2: ???

## Glass Data

Reading in the data from the previous homework:

```{r}
glass_names = c("ID","RI","NA","MG","AL","SI","K","CA","BA","FE","Type_of_glass")
glass_Data <- read_csv("https://www4.stat.ncsu.edu/online/datasets/glass.data",col_names = glass_names)
```
Creating the factor variable and filtering output:

```{r}
glass_data_filtered <- glass_Data |>
  mutate(Type_Of_Glass = factor(
    Type_of_glass,
    levels = 1:7,
    labels = c(
      "building_windows_float_processed",
      "building_windows_non_float_processed",
      "vehicle_windows_float_processed",
      "vehicle_windows_non_float_processed",
      "containers",
      "tableware",
      "headlamps"
    )
  )) |>
  filter(FE<0.2,Type_Of_Glass %in% c("tableware","headlamps"))

```


## Yeast Data

Reading in data: 

```{r}
yeast_names <- c("seq_name","mcg","gvh","alm","mit","erl","pox","vac","nuc","class")
yeast_data <- read_table("https://www4.stat.ncsu.edu/online/datasets/yeast.data",col_names = yeast_names)
```

Removing columns and adding mean/median columns:

```{r}
yeast_data_summary <- yeast_data |>
  select(-1,-9) |>
  group_by(class)|>
  mutate(across(1:7,list(mean=mean,median=median),.names="{.col}_{.fn}"))
```

## Wine Data

All of the reading in from previous assignment:

```{r}
library(readxl)
white_wine <- read_excel("data/white-wine.xlsx",col_types = "numeric")
wine_names <-read_excel("data/white-wine.xlsx", sheet = excel_sheets("data/white-wine.xlsx")[2])
colnames(white_wine) <- wine_names$Variables
white_wine$type <- "white"
red_wine <- read_delim("https://www4.stat.ncsu.edu/~online/datasets/red-wine.csv",delim = ";")
colnames(red_wine) <- wine_names$Variables
red_wine$type <- "red"
wine_data <- bind_rows(white_wine,red_wine)
```
Filtering the wine data:

```{r}
wine_data_filtered <- wine_data |>
  filter(quality > 6.5, alcohol < 132) |>
  arrange(desc(quality))|>
  select(contains("acid"),alcohol,type,quality)|>
  group_by(quality)|>
  mutate(across(-5,list(mean=mean,median=median),.names="{.col}_{.fn}"))
```













