---
title: "Homework 5"
author: "Carlos Petzold"
format: html
---

### Task 1 - Querying an API and a Tidy-Style Function

#### 1) Use GET() from the httr package to return information about a topic that you are interested in that has been in the news lately (store the result as an R object). Note: We can only look 30 days into the past with a free account.

```{r Task 1 Question 1}

#pull needed libraries
library(httr)
library(jsonlite)

#api key
news_api_key <- "0d36c91e11824ce488ef692db17fe779"

#pull news
data <- GET(paste0("https://newsapi.org/v2/everything?q=Dortmund",
                   "&from=2026-09-14",
                   "&language=en",
                   "&apiKey=", news_api_key))

#check that it was successful
status_code(data)


```

#### 2) Parse what is returned and find your way to the data frame that has the actual article information in it (check content). Note the first column should be a list column!

```{r Task 1 Question 2}

#pull content
soccer_news_content <- content(data, as = "text", encoding = "UTF-8") |>
  fromJSON()

str(soccer_news_content, max.level = 1)

# Store the article information as a df
articles <- soccer_news_content$articles

articles

```

#### 3) Now write a quick function that allows the user to easily query this API. The inputs to the function should be the title/subject to search for (string), a time period to search from (string - you’ll search from that time until the present), and an API key.

```{r Task 1 Question 3}

 get_news <- function(subject, date, api_key) {
   
   url <- paste0("https://newsapi.org/v2/everything?q=", subject,
                 "&from=", date,
                 "&language=en",
                 "&apiKey=", api_key)
   
   result <- GET(url)
   parsed <- content(result, as = "text", encoding = "UTF-8") |>
     fromJSON()
   
   return(parsed$articles)
   
 }

#Articles about Bundesliga:
bundesliga_news <- get_news("Bundesliga", "2026-09-14", news_api_key)

#Articles about Fifa:
NFL_news <- get_news("NFL", "2026-09-14", news_api_key)

```

