---
title: "Homework 2"
author: "Group 8"
format: html
---

## Task 1: Basic Vector Practice 

### 1. Creating the Blood Pressure Vectors

Two vectors are created corresponding to the pre-treatment and post-treatment measurements for twenty subjects.

```{r Task 1 Question 1}

#Pre Measurements
pre <- c(130, 128, 116, 124, 133, 134, 118, 126, 114, 127, 141, 138, 128, 140, 137, 131, 120, 128, 139, 135)

#Post Measurements
post <- c(114, 98, 113, 99, 107, 116, 113, 111, 119, 117, 101, 119, 130, 122, 106, 106, 124, 102, 117, 113)

```

### 2. Assigning Names to the Vector Elements
Subject names are assigned using the paste() function to the pre-treatment and post-treatment vectors.


```{r Task 1 Question 2}

#Assigning names to the vector elements (1st element becomes Subject_1_Pre, and so on)
names(pre) <- paste("Subject", 1:20, sep = "_")

#Assigning names to the vector elements (1st element becomes Subject_1_Post, and so on)
names(post) <- paste("Subject", 1:20, sep = "_")

#Print names
names(pre)
names(post)
```

### 3. Calculating Change in Blood Pressure for Each Subject

Calculated the change in blood pressure (pretreatment - posttreatment) and saved the calculation as a new vector.

```{r Task 1 Question 3}

bp_diff = pre - post

bp_diff

```

### 4. Calculating Average Decrease For All Subjects

Calculated the average decrease in blood pressure across all subjects.

```{r Task 1 Question 4}

mean(bp_diff)

```

### 5. Determining which Subjects Experienced Decrease in BP with Treatment

Determined a list of subjects that experienced a decrease in blood pressure (diff_bp\>0) with treatment.

```{r Task 1 Question 5}

which(bp_diff > 0)

```

### 6. Subset Vectors

Subset the vector of differences to only return those that have a positive change.

```{r Task 1 Question 6}

bp_diff_subset <- bp_diff[bp_diff > 0]

bp_diff_subset

```

### 7. Calculating Average Decrease for Subset of Subjects

Calculated the average decrease in blood pressure for those subjects that had a positive change

```{r Task 1 Question 7}

mean(bp_diff_subset)

```


## Task 2: Basic Data Frame Practice

### 1. Creating a Data Frame

Created a data frame object with four columns corresponding to data from Task 1.

```{r Task 2 Question 1}

bp_df <- data.frame(patient = 1:20, pre_bp = pre, post_bp = post, diff_bp = bp_diff)

bp_df

```

### 2. Returning Rows with Select Column Values

Returned only rows where diff_bp column is negative by accessing the column from data frame.


```{r Task 2 Question 2}

#referencing the diff_bp column of the bp_df data frame
bp_df[bp_df$diff_bp < 0, ]

```

### 3. Adding a Column to Existing Data Frame

Added a column to the data frame corresponding to "TRUE" if post_bp \< 120

```{r Task 2 Question 3}

bp_df$bp_lt_120 <- bp_df$post_bp < 120

bp_df

```

### 4. Printing Data Frames

Displayed completed blood pressure data frame in a formatted table.

```{r Task 2 Question 4}

knitr::kable(bp_df)

```


## Task 3: List Practice

### 1. Creating Placebo Data Frame

Pre-treatment and post-treatment blood pressure measurements from the placebo experiment are used to create a second data frame similar to the treatment data frame.

```{r Task 3 Question 1}

#creating new data frame called new_df
new_df <- data.frame(
  patient = 1:10,
  pre_bp = c(138, 135, 147, 117, 152, 134, 114, 121, 131, 130), # Create placebo pre-treatment blood pressure vector
  post_bp = c(105, 136, 123, 130, 134, 143, 135, 139, 120, 124) # Create placebo post-treatment blood pressure vector
)

# Calculate change in blood pressure for placebo
new_df$diff_bp <- new_df$pre_bp - new_df$post_bp

# Add a TRUE/FALSE column based on post-placebo blood pressure
new_df$bp_lt_120 <- new_df$post_bp < 120

new_df
```

### 2. Create and Store Lists

A list is created to store the treatment and placebo data frames as separate named elements.

```{r Task 3 Question 2}

#creating a new list called new_list which contains our treatment and placebo datasets
new_list <- list(treatment = bp_df, placebo = new_df)

new_list

```

### 3. Accessing First List Element

The treatment data frame, which is the first element of "new_list", is accessed using three different list-indexing methods.

```{r Task 3 Question 3}

# Access first element using double brackets
new_list[[1]]

# Access first element using its name
new_list$treatment

# Access first element using single brackets
new_list[1]

```

### 4. Accessing a Specified Column of a Data Frame

The "pre_bp" column is accessed directly from the placebo data frame stored within "new_list".

```{r Task 3 Question 4}

# Access the placebo pre-treatment blood pressure column
new_list[[2]]$pre_bp

```


## Task 4: Control Flow Practice

### 1. Creating Status Column

A new character column called "status" is added to each data frame stored in "new_list". These columns will later be filled using conditional logic based on post-treatment blood pressure. 

```{r Task 4 Question 1}

new_list[[1]]$status <- character(20)

new_list[[2]]$status <- character(10)

```

### 2. Assigning Status for Treatment Group

A "for loop" and conditional statements are used to assign blood pressure status to each treatment subject based on their post-treatment blood pressure.

```{r Task 4 Question 2}

#looping through the full length of the status variable from the first element of the new_list list, which is the non-placebo data frame
for (i in 1:20) {
  
  if (new_list[[1]]$post_bp[i] <= 120) {
    
    new_list[[1]]$status[i] <- "Optimal"
    
  } else if (new_list[[1]]$post_bp[i] > 120 & new_list[[1]]$post_bp[i] <= 130) {
    
    new_list[[1]]$status[i] <- "Borderline"
    
  } else {
    
    new_list[[1]]$status[i] <- "High"
    
  }
  
}

new_list$treatment

```

### 3. Assigning Status for Placebo Group

A "for loop" and conditional statements are used to assign blood pressure status to each placebo subject based on their post-treatment blood pressure.

```{r Task 4 Question 3}

#looping through the full length of the status variable from the second element of the new_list list, which is the placebo data frame
for (i in 1:10) {
  
  if (new_list[[2]]$post_bp[i] <= 120) {
    
    new_list[[2]]$status[i] <- "Optimal"
    
  } else if (new_list[[2]]$post_bp[i] > 120 & new_list[[2]]$post_bp[i] <= 130) {
    
    new_list[[2]]$status[i] <- "Borderline"
    
  } else {
    
    new_list[[2]]$status[i] <- "High"
    
  }
  
}

new_list$placebo

```


