---
title: "ST558_Homework3_Sheth"
format: html
editor: visual
---

## Task 1: Conceptual Questions

1.  **If your working directory is myfolder/homework/, what relative path would you specify to get the file located at myfolder/MyData.csv?** The relative path for this file is "../MyData.csv", where the ".." indicates moving up one directory from the current "homework" folder before accessing "MyData.csv".

2.  **What are the major benefits of using R projects?** R projects help keep analyses organized by giving each project its own working directory, workspace, history, folder structure, and source documents. They also make it easier to use relative file paths, which improves reproducibility and facilitates collaboration by enabling running the same code on different computers/by different users.

3.  **What is Git and what is GitHub?** Git is a version control system that tracks changes made to files over time. It allows users to save versions of their work, review changes, and collaborate without losing previous versions. GitHub is an online platform that hosts Git repositories. It allows users to store their projects remotely, share them with others, collaborate on the same files, and manage changes through tools such as commits, branches, and pull requests.

4.  **What are the two main differences between a tibble and a data.frame?** A tibble is a special type of data frame. It has additional classes (eg, tbl_df and tbl) while still inheriting from data.frame, so it can generally be used anywhere a regular data frame can be used. Additionally, tibble has a different default printing behavior, where when printed it shows the dimensions, variable types, and only rows/columns that fit reasonably on the screen rather than flooding the console with the entire output.

## Task 2: Reading Delimited Data

Loading "tidyverse" to support reading in data using readr function. 

```{r}
# Load tidyverse
library(tidyverse)
```

### 1. Reading the Glass Data
The glass data are comma delimited data and do not contain column names. The data are read directly from the provided URL using the readr function in the tidyverse library, and column names are assigned at the time of importing the data set. The tibble is printed out by calling the object name (i.e., "glass_data"). 

```{r}
# Assign column names for data set
glass_names <- c("Id", "RI", "Na", "Mg", "Al", "Si", "K", "Ca", "Ba", "Fe", "Type")

# Read in comma delimited data from URL
glass_data <- read_csv("https://www4.stat.ncsu.edu/~online/datasets/glass.data", 
                       col_names = glass_names
                      )
# Print the tibble
glass_data
```

### 2. Reading the Yeast Data
The yeast data are space (or whitespace) delimited data and do not contain column names. The data are read directly from the provided URL using the readr function in the tidyverse library, and column names are assigned at the time of importing the data set. The tibble is printed out by calling the object name (i.e., "yeast_data").

```{r}
# Assign column names for the data set
yeast_names <- c("seq_name", "mcg", "gvh", "alm", "mit", "erl", "pox", "vac", "nuc", "class")

# Read in space (whitespace) delimited data from URL
yeast_data <- read_table("https://www4.stat.ncsu.edu/~online/datasets/yeast.data",
                         col_names = yeast_names
                         )
# Print the tibble
yeast_data
```

## Task 3: Combining Excel and Delimited Data

### 3. Reading the White Wine Excel Data
The white wine data are stored in an Excel workbook with column names. The first sheet is imported using the "readxl" package and a relative path to the file stored in the project's "data" folder.

```{r}
# Load readxl
library(readxl)

# Read the first sheet of the white wine Excel file
white_wine <- read_excel("data/white-wine.xlsx",
                         sheet = 1
                         )

# Print the tibble
white_wine
```
### 4. Replacing the White Wine Variable Names
The second sheet of the Excel workbook contains alternative variable names without spaces. These names are read into R and used to replace the current names in the white wine tibble. 

```{r}
# Read in the second sheet containing alternative variable names
wine_names <- read_excel("data/white-wine.xlsx",
                         sheet = 2
                         )
wine_names

# Replace the current white wine column names
colnames(white_wine) <- wine_names[[1]]

# Print the updated tibble
white_wine
```
### 5. Adding the Wine Type
A new variable named "type" is added to the white wine tibble to identify each observation as white wine. The value "white" is assigned to each row. 

```{r}
# Add a column identifying the wine type
white_wine$type <- "white"

# Print the updated tibble
white_wine
```

### 6. Reading the Red Wine CSV Data
The red wine data are stored in a semicolon delimited file. The data are read directly from the provided URL using the readr function. The variable names are then replaced using the alternative names from the Excel file for white wine. Additionally, a variable is added to identify each observation as red wine. 

```{r}
# Read in semicolon delimited data
red_wine <- read_delim("https://www4.stat.ncsu.edu/~online/datasets/red-wine.csv",
                       delim = ";"
                       )

# Replace the red wine column names
colnames(red_wine) <- wine_names[[1]]

# Add a column identifying the wine type
red_wine$type <- "red"

# Print the tibble
red_wine
```
### 7. Combining Data Sets
The white wine and red wine tibbles have the same variables, so they are combined by rows using "dplyr::bind_rows()".

```{r}
# Combine the white and red wine data sets
wine_data <- bind_rows(white_wine, red_wine)

# Print the combined tibble
wine_data
```


## Task 4: Writing a Function
