---
title: "Homework 3"
author: "Carlos Petzold"
format: html
---

# Task 1: Conceptual Questions

* 1) If your working directory is myfolder/homework/, what relative path would you specify to get the file located at myfolder/MyData.csv?
 
  * If my working directory is myfolder/homework/, the relative path that I specify to get the file located at myfolder/Mydata.csv is ../MyData.csv

* 2) What are the major benefits of using R projects?

  * The major benefits of using R Projects are reproducibility and the ability to easily work with others on the same set of files. Each R project has its own working directory, workspace, history, and folder structure, so by sharing an R project with another person, thet will be able to pick it up and use the paths without any headache. 
  
* 3) What is Git and what is GitHub?

  * Git is a version control software that tracks changes made to files that are being used my multiple people, and GitHub is a web platform where programmers can share and manage their programs/code. 
  
* 4) What are the two main differences between a tibble and a data.frame?

  * A tibble does not print completely in the console when called, but a data frame does. Also, tibbles do not coerce down to a vector when you subset to only 1 column using [].
  
  
# Task 2: Reading Delimited Data

## First, we are reading in the glass.data dataset from the UCI machine learning repository

```{r Task 2, Glass Data}

#We'll use read_csv from the readr package to read in the csv file from the URL:

glass_data <- readr::read_csv("https://www4.stat.ncsu.edu/~online/datasets/glass.data",
                       col_names = c("ID", "refractive_index", "sodium",
                                     "magnesium", "aluminum", "silicon",
                                     "potassium", "calcium", "barium",
                                     "iron", "glass_type"))

glass_data

```


## Next, we are reading in the yeast.data dataset from the UCI machine learning repository

```{r Task 2, Glass Data}

#We'll use read_tsv from the readr package to read in the tab-seperated file from the URL:

glass_data <- readr::read_table("https://www4.stat.ncsu.edu/~online/datasets/yeast.data",
                         col_names = c("seq_name", "mcg", "gvh", "alm", "mit", "erl",
                                       "pox", "vac", "nuc", "class"))

glass_data

```


# Task 3: Combining Excel and Delimited Data

## First we will read in the first sheet of the white wine excel file

```{r Task 3, read Wine data first sheet}

#Reading in first sheet of the file
wine_data <- readxl::read_excel("data/white-wine.xlsx", sheet = 1)

wine_data

```

## Next we will read in the second sheet of the white wine excel file, and assign

```{r Task 3, read Wine data second sheet}

#Reading in second sheet of the file
wine_data_colnames <- readxl::read_excel("data/white-wine.xlsx", sheet = 2)

wine_data_colnames

#Assigning the vector of names from the second sheet to be the column names of the first sheet
colnames(wine_data) <- wine_data_colnames[[1]]

```

## Next we will add a column to indicate if the wines are white (in this case, all are)

```{r Task 3, Wine data white indicator}

#Using dplyr::mutate to add a white wine indicator
wine_data <- wine_data |>
  dplyr::mutate(wine_type = "white")

```

## Next we will read in a red wine dataset (semi-colon delimited)

```{r Task 3, Wine data white indicator}

#Using dplyr::mutate to add a white indicator
red_wine_data <- readr::read_delim("https://www4.stat.ncsu.edu/~online/datasets/red-wine.csv",
                                  delim = ";")

#replacing column names
colnames(red_wine_data) <- wine_data_colnames[[1]]

#Using dplyr::mutate to add a red wine indicator
red_wine_data <- red_wine_data |>
  dplyr::mutate(wine_type = "red")

```

## Lastly, we will combine the 2 wine data sets

```{r Task 3, Wine data white indicator}

#combining the 2 datasets using dplyr::bind_rows
wine_data_combined <- dplyr::bind_rows(wine_data, red_wine_data)

```

