---
title: "Homework 4"
author: "Carlos Petzold"
format: html
---

## Task 1: Conceptual Questions

### 1) Rewrite the following nested function call using baseR’s chaining operator:

library(tidyverse)
arrange(filter(select(as_tibble(iris), starts_with("Petal"), Species), Petal.Length <
1.55), Species)


```{r Task 1 Question 1}

library(tidyverse)

iris <- as_tibble(iris) |> 
  select(starts_with("Petal"), Species) |>
  filter(Petal.Length < 1.55) |>
  arrange(Species)

```


### 2) What is meant by long format data and wide format data? Which do we generally prefer for statistical analysis?

Long data format refers to data that is formatted with 1 observation per row and each column corresponds to one variable. Wide data format refers to data that may have more than one observation for a given row. When conductin statistical analysis, we typically prefer data in the long format because it makes plotting, summarizing, modeling, and analyzing data much simpler when we have one observation per row.


## Task 2: Read in Datasets

### 3) Glass Data: Start a chain on your data object that overwrites the Type_of_glass variable using mutate().

#### Create a factor version of the variable that replaces the numbers with string labels (that is, replace 1 with “building_windows_float_processed”, 2 with “building_win. . . ”, etc. - see the variable descriptions above to give meaningful values).

### 4) Glass Data: Continue your chain and keep only observations where the Fe variable is less than 0.2 and the Type of Glass is either “tableware” or “headlamp”.

```{r Task 2 Question 3 and 4}

#reading in glass data as a csv
glass_data <- readr::read_csv("https://www4.stat.ncsu.edu/~online/datasets/glass.data",
                       col_names = c("ID", "refractive_index", "sodium",
                                     "magnesium", "aluminum", "silicon",
                                     "potassium", "calcium", "barium",
                                     "Fe", "type_of_glass")) |>
  mutate(type_of_glass = factor(
    glass_type,
    levels = c(1, 2, 3, 4, 5, 6, 7),
    labels = c(
      "building_windows_float_processed",
      "building_windows_non_float_processed",
      "vehicle_windows_float_processed",
      "vehicle_windows_non_float_processed",
      "containers",
      "tableware",
      "headlamps"
    )
  )) |>
  filter(Fe < 0.2, type_of_glass %in% c("tableware", "headlamp"))
  
glass_data

```


### 5) Yeast Data: Start a chain on your data object that overwrites the Type_of_glass variable using mutate().

### 6) Yeast Data: Continue your chain to add columns corresponding to the mean and median of each numeric variable (mcg, gvh, alm, mit, erl, pox, and vac) at each class grouping (see the across() function as we did in the dplyr video!).

```{r Task 2 Question 5 and 6}

#reading in yeast data using read_table
yeast_data <- readr::read_table("https://www4.stat.ncsu.edu/~online/datasets/yeast.data",
                         col_names = c("seq_name", "mcg", "gvh", "alm", "mit", "erl",
                                       "pox", "vac", "nuc", "class")) |>
  select(-seq_name, -nuc) |>
  group_by(class) |>
  mutate(across(where(is.numeric), \(x) mean(x), .names = "mean_{.col}"),
         across(where(is.numeric), \(x) median(x), .names = "median_{.col}"))

yeast_data

```


### 7) Wine Data: Start a chain on your new combined data object to filter the data to only look at high-quality wines (quality > 6.5) and wines that have a reasonable alcohol value (alcohol < 13.2).

### 8) Wine Data: Continue your chain to now sort the data from highest quality to lowest.

### 9) Wine Data: Continue your chain to select only the variables that contain acid, the alcohol variable, the type variable, and the quality variable

### 10) Wine Data: Continue your chain to add the mean and standard deviation of the alcohol variable to the data set for each setting of the quality variable.

```{r Task 2 Question 7, 8, 9, and 10}

#Reading in first and second sheet of the file
white_wine_data <- readxl::read_excel("data/white-wine.xlsx", sheet = 1)
wine_data_colnames <- readxl::read_excel("data/white-wine.xlsx", sheet = 2)

#Assigning the vector of names from the second sheet to be the column names of the first sheet
colnames(white_wine_data) <- wine_data_colnames[[1]]

#Using dplyr::mutate to add a white wine indicator
white_wine_data <- white_wine_data |>
  dplyr::mutate(type = "white")

#Reading in red wine data
red_wine_data <- readr::read_delim("https://www4.stat.ncsu.edu/~online/datasets/red-wine.csv",
                                  delim = ";")

#replacing column names
colnames(red_wine_data) <- wine_data_colnames[[1]]

#Using dplyr::mutate to add a red wine indicator
red_wine_data <- red_wine_data |>
  dplyr::mutate(type = "red")

#combining the 2 datasets using dplyr::bind_rows
wine_data_combined <- dplyr::bind_rows(white_wine_data, red_wine_data)

wine_data <- wine_data_combined |>
  filter(quality > 6.5, alcohol < 13.2) |>
  arrange(desc(quality)) |>
  select(fixed_acidity, volatile_acidity, citric_acid, alcohol, type, quality) |>
  group_by(quality) |>
  mutate(mean_alcohol = mean(alcohol),
         sd_alcohol = sd(alcohol))
  

```


## Task 3: Database Practice

#### Download the Lahman.db file associated with the Connecting to Databases notes/video.

### 11) Connect to the database and then look at all of the tables in the database.

```{r Task 3 Question 11}

con <- DBI::dbConnect(RSQLite::SQLite(), "data/lahman.db")
DBI::dbListTables(con)

```

### 12) Use the tbl() function and dplyr to return all of the data from the Teams table for the year 2015.

```{r Task 3 Question 12}

teams_2015 <- dplyr::tbl(con, "Teams") |> 
  filter(yearID == 2015)

teams_2015

```

### 13) Repeat the above by using the sql() function within tbl() (here you have to write actual SQL!).

```{r Task 3 Question 13}

dplyr::tbl(con, "Teams") |> 
  filter(yearID == 2015) |>
  show_query()

teams_2015_sql <- tbl(con, sql(
  "SELECT `Teams`.*
    FROM `Teams`
    WHERE (`yearID` = 2015.0)"
))

teams_2015_sql

```

### 14) Return all of the players in the hall of fame, the year they were voted into the hall of fame, and their category (only those three variables!). See the HallOfFame table, the inducted variable is important here.

```{r Task 3 Question 14}

hof_players <- dplyr::tbl(con, "HallOfFame") |>
  filter(inducted == "Y") |>
  select(playerID, yearID, category) 

hof_players

```

### 15) Combine the table from the previous question with the People table in order to have the nameFirst and nameLast names added to what was returned above.

```{r Task 3 Question 15}

hof_players_w_names <- hof_players |>
  left_join(dplyr::tbl(con, "People") |>
              select(playerID, nameFirst, nameLast),
            by = "playerID") |>
  collect()

hof_players_w_names

```

### 16) Return only the playerID (manager ID in this case), G, W, and L columns from the Managers table. Use chaining to then:

#### Determine the overall win/loss records (sum of wins and sum of losses) for each of these hall of fame managers.

#### Create a new variable that is the career win/loss percentage (Total_W/G_managed). (Note: I did this after collect() otherwise the column type has to be set. . . )

#### Sort the resulting data by the win/loss percentage variable (from largest to smallest).

```{r Task 3 Question 16}

managers <- dplyr::tbl(con, "Managers") |>
  select(playerID, G, W, L) |>
  group_by(playerID) |>
  summarize(G_managed = sum(G, na.rm = TRUE),
            Total_W = sum(W, na.rm = TRUE),
            Total_L = sum(L, na.rm = TRUE)) |>
  ungroup() |>
  collect() |>
  mutate(w_l_percentage = Total_W/G_managed) |>
  arrange(desc(w_l_percentage))
  

managers
```

### 17) Use the results from the previous two questions to answer this one! Return information for only the people that managed a team and were inducted into the hall of fame (regardless of their category in the hall of fame - you want to return all people in the hall of fame that ever managed at all). (Just return all of the variables from both of the above two questions - 9 total columns):

```{r Task 3 Question 17}

hof_managers <- inner_join(managers, as.tibble(hof_players_w_names), by = "playerID")

hof_managers

```




