---
title: "ST558_Homework3_King"
format: html
---

# Task 1: Conceptual Questions

1.  The working path for this file would be "../MyData.csv"
2.  The major benefits of using R Projects are the ability to manage multiple files associated with a project, and the integration with GitHub.
3.  Git is version control software, and GitHub is an online hosting service for Git projects.
4.  Tibbles differ from data frames in that they print more nicely, and when you subset one column of a tibble, it doesn't turn into a vector, it stays a tibble.

# Task 2: Reading Delimited Data

## Glass Data

```{r}
# loading in tidyverse
library(tidyverse)
# the raw data is comma delimited, so we use read_csv. the columns names are set to a vector of the names.
glassData <- read_csv("https://www4.stat.ncsu.edu/online/datasets/glass.data",col_names = c("ID","RI","NA","MG","AL","SI","K","CA","BA","FE","Type"))
# print the tibble
glassData
```

## Yeast Data

```{r}
# the raw data is space delimited, so we use read_table.
yeastData <- read_table("https://www4.stat.ncsu.edu/online/datasets/yeast.data",col_names = c("seq_name","mcg","gvh","alm","mit","erl","pox","vac","nuc","class"))
# print the tibble
yeastData
```

# Task 3: Combining Excel and Delimited Data

## Initially Reading the Sheet

```{r}
# loading in readxl
library(readxl)
# read in data from the first sheet using read_excel
whiteData <- read_excel("data/white-wine.xlsx",col_types = "numeric")
whiteData
```

## Changing the Column Names
```{r}
whiteVars <-read_excel("data/white-wine.xlsx", sheet = excel_sheets("data/white-wine.xlsx")[2])
colnames(whiteData) <- whiteVars$Variables
```

## Adding the White Wine Column
```{r}
whiteData$type <- "white"
```

## Reading in Red Wine

```{r}
# repeating the process of reading delimited data, but with semicolons this time
redData <- read_delim("https://www4.stat.ncsu.edu/~online/datasets/red-wine.csv",col_types = "nnnnnnnnnnnn",delim = ";")
# changing variable names
colnames(redData) <- whiteVars$Variables
# adding the type column
redData$type <- "red"
```

## Combining White and Red
```{r}
wineData <- bind_rows(whiteData,redData)
wineData
```

# Task 4: Writing a Function

## Writing getIQR Function

```{r}
getIQR <- function(x){
  as.numeric(quantile(x,.75)-quantile(x,.25))
}
```

## Applying getIQR to fixed_acidity and citric_acid

```{r}
getIQR(wineData$fixed_acidity)
getIQR(wineData$citric_acid)
```
## Writing getCoefVar Function

```{r}
test<-1:10
sd(test)/abs(mean(test))

getCoefVar(test)

getCoefVar <- function(x){
  sd(x)/abs(mean(x))
}


```


















