---
title: "zhking_hw2"
format: html
---

Note: I haven't looked at Aliyah's version, but I don't think there's anything wrong with it. I just want to show that I have also done the homework.

# Task One: Basic Vector Practice

1. Establishing pre- and post- treatment vectors.

```{r}
pre <- c(130, 128, 116, 124, 133, 134, 118, 126, 114, 127, 141, 138, 128, 140, 137, 131, 120, 128, 139, 135)
post <- c(114, 98, 113, 99, 107, 116, 113, 111, 119, 117, 101, 119, 130, 122, 106, 106, 124, 102, 117, 113)
```

2. Assign names to the vector elements

```{r}
names(pre) <- paste("Subject", 1:20, sep = "_")
names(post) <- paste("Subject", 1:20, sep = "_")
```

3. Calculate the change in blood pressure

```{r}
change <- pre-post
```

4. Calculate average decrease

```{r}
mean(change)
```
5. Find subjects with a decrease in blood pressure

```{r}
which(change>0)
```
6. Subset vector of only decreases

```{r}
decreases <- change[which(change>0)]
```

7. Average decrease in those who decreased

```{r}
mean(decreases)
```

# Task 2: Basic Data Frame Practice

1. Create a data frame with previous vectors

```{r}
data <- data.frame(patient=names(pre),pre_bp=pre,post_bp=post,diff_bp=change)
```

2. Return only rows where change in bp is negative

```{r}
data[which(data$diff_bp<0),]
```

3. Add a new column to the data frame that is TRUE when post_bp is < 120

```{r}
data$good_bp <- data$post_bp < 120
```

4. Print Nicely
```{r}
knitr::kable(data)
```

# Task 3: List Practice

1. 

```{r}
pre_placebo <- c(138, 135, 147, 117, 152, 134, 114, 121, 131, 130)
post_placebo <- c(105, 136, 123, 130, 134, 143, 135, 139, 120, 124)
names(pre_placebo) <- paste("Subject", 1:10, sep = "_")
placebo_bp <- data.frame(pre_placebo,post_placebo,diff_placebo=(pre_placebo-post_placebo),good_bp=post_placebo<120)
```


2. Creating a List of the Dataframes

```{r}
bp_data_list <- list(treatment=data,placebo=placebo_bp)
```

3. Accessing First List Element

```{r}
bp_data_list$treatment
bp_data_list[1]
bp_data_list[[1]]
```

4. Accessing the Second Column of the First Element of the List

```{r}
bp_data_list$treatment[,2]
```

